### 2018-01-23 notes/2018-01-18-osa-abstract/tester.tex

\documentclass{article}
\usepackage[demo]{graphicx}
\usepackage{floatrow}
\usepackage[export]{adjustbox}

\begin{document}

\begin{figure}
    \includegraphics[width=1\textwidth,left]{name}
    \caption{centered image}
\end{figure}

\begin{figure}
\flushleft
\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=6cm}}]{figure}[\FBwidth]
{\caption{A test figure with its caption side by side}\label{fig:test}}
{\includegraphics[width=4cm]{name}}
\end{figure}

\begin{figure}
\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=4cm}}]{figure}[\FBwidth]
{\caption{A test figure with its caption side by side}\label{fig:test}}
{\includegraphics[width=5cm]{name}}
\end{figure}

\end{document}