### 2018-03-12 slides/2018-03-12-group-meeting/group_meeting.tex
Group meeting slides w/ group theory.

\documentclass[presentation]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\tolerance=1000
\usepackage{graphicx} \DeclareMathOperator{\argmin}{argmin}

\newcommand{\me}{\mathrm{e}}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mf}[1]{\mathfrak{#1}}
\providecommand{\ro}[1]{\mathbf{\mathfrak{r}}_o}
\providecommand{\so}[1]{\mathbf{\hat{s}}_o}
\providecommand{\rb}[1]{\mathbf{r}_b}
\providecommand{\rbm}[1]{r_b^{\text{m}}}
\providecommand{\rd}[1]{\mathbf{r}_d}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\bs}[1]{\boldsymbol{#1}} 
\providecommand{\intinf}{\int_{-\infty}^{\infty}}

\newcommand{\under}[2]{\underset{\scriptscriptstyle#1}{#2}}

\usetheme{simple}
\usecolortheme{}
\usefonttheme{serif}
\useinnertheme{}
\useoutertheme{}
\author{Talon Chandler}
\date{March 12, 2018}
\title{Update on Spatio-angular Transfer Functions}

\begin{document}

\maketitle
\begin{frame}[label=sec-1]{Steady progress on closed form OTF}
  \begin{itemize}
  \item Initial calculation didn't match my numerical OTF.
  \item Found some issues that I glossed over. Calculation still in progress.
  \end{itemize}
\end{frame}

\begin{frame}[label=sec-1]{Brute-force degeneracy finding}
 \begin{center}
   \includegraphics[width=0.8\textwidth]{figs/solid-angle-uncertainty.png}
 \end{center}
\end{frame}

\begin{frame}[label=sec-1]{0.8 NA Epi-illumination Microscope}
 \begin{center}
   \includegraphics[width=0.8\textwidth]{figs/singular-value-spectrum-epi.pdf}
 \end{center}
\end{frame}

\begin{frame}[label=sec-1]{0.8 NA Symmetric diSPIM Microscope}
 \begin{center}
   \includegraphics[width=0.8\textwidth]{figs/singular-value-spectrum-dispim.pdf}
 \end{center}
\end{frame}

\begin{frame}[label=sec-1]{Group theory for finding degeneracy}
\begin{columns}
  \begin{column}{0.5\textwidth}
    \centering \textbf{Quantum mechanics}\\\vspace{2em}
    $\mathcal{H} = \text{Hamiltonian}$\\\vspace{2em}
    $\mathcal{H}\psi_k = W_k\psi_k$\\\vspace{2em}
    $\mathcal{T}_j\mathcal{H} - \mathcal{H}\mathcal{T}_j = 0$\\\vspace{2em}
    $\mathcal{T}_j$ form the symmetry group of the Hamiltonian.
\end{column}
\begin{column}{0.5\textwidth}  %%<--- here
  \centering \textbf{Imaging}\\\vspace{2em}
  $\mathcal{H} = \text{Forward operator}$\vspace{2em}
  $\mathcal{H}^{\dagger}\mathcal{H}u_k = \mu_ku_k$\\\vspace{2em}
  $\mathcal{T}_j\mathcal{H}^{\dagger}\mathcal{H} - \mathcal{H}^{\dagger}\mathcal{H}\mathcal{T}_j = 0$\\\vspace{2em}
  $\mathcal{T}_j$ form the symmetry group of the imaging system.

\end{column}
\end{columns}
\end{frame}

\end{document}