### 2018-11-07 papers/basic-theory/figures/dipole-block/dipole-block.tex
Add latest figs and bib.

\documentclass[crop,tikz]{standalone}% 'crop' is the default for v1.0, before it was 'preview'
% \usetikzlibrary{...}% tikz package already loaded by 'tikz' option
\usepackage{tikz}
\usepackage{amsmath, amsthm, amssymb, amsfonts}
\usetikzlibrary{shapes,arrows.meta}
\usepackage{newtxtext,newtxmath,courier}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}

% Talon's custom commands
\usepackage{xcolor}
\newcommand{\gray}[1]{\colorbox{black!10!white}{$\displaystyle #1\hspace{-0.2em}$}\,}
\newcommand{\argmin}{\arg\!\min}
\newcommand{\me}{\mathrm{e}}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mf}[1]{\mathfrak{#1}}
\providecommand{\mc}[1]{\mathcal{#1}}
\providecommand{\ro}{\mathbf{\mathbf{r}}_o}
\providecommand{\so}{\mathbf{\hat{s}}_o}
\providecommand{\rb}{\mathbf{r}_b}
\providecommand{\rbm}[1]{r_b^{\text{m}}}
\providecommand{\rd}{\mathbf{r}_d}
\providecommand{\rdf}{\mathpzc{r}_d}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\mbb}[1]{\mathbb{#1}}
\providecommand{\bs}[1]{\boldsymbol{#1}}
\providecommand{\bv}{\bs{\nu}}
\providecommand{\bsh}[1]{\hat{\boldsymbol{#1}}}
\providecommand{\nan}{\left(\frac{\text{NA}}{n_o}\right)}
\providecommand{\lmsum}{\sum_{l=0}^\infty\sum_{m=-l}^{l}}
\providecommand{\intr}[1]{\int_{\mbb{R}^{#1}}}
\providecommand{\ints}[1]{\int_{\mbb{S}^{#1}}}
\DeclareFontFamily{OT1}{pzc}{}
\DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\newcommand{\eqname}[1]{\tag*{#1}}
\newcommand*\widefbox[1]{\fbox{\hspace{1em}#1\hspace{1em}}}

\tikzstyle{block} = [draw, fill=white, rectangle, minimum height=5em, minimum width=7em, align=center]

\begin{document}
  \begin{tikzpicture}[auto, node distance=2cm,>={Latex[length=2mm]}]
    \tikzset{font={\fontsize{9pt}{12}\selectfont}}
    \node [block, align=center] (sad) {$f(\ro{}, \so)$\\ Dipole\\ density};
    \node [block,  right of=sad, node distance=4cm] (ads) {$F_\ell^m(\ro)$\\ Angular\\ dipole\\ spectrum};
    \node [block,  right of=ads, node distance=9cm] (irr) {$g(\rd)$\\ Scaled\\ irradiance};
    \node [block, below of=ads, node distance=2.5cm] (sas) {$\mathsf{F}_\ell^m(\bs{\nu})$\\ Spatio-angular\\ dipole\\ spectrum};
    \node [block,  right of=sas, node distance=9cm] (is) {$G(\bs{\nu})$\\ Scaled\\ irradiance\\ spectrum};
    \node [block, below of=sad, node distance=2.5cm] (sds) {$F(\bs{\nu}, \so)$\\ Spatial\\ dipole\\ spectrum};
    \draw [->] (sad) -- node[] {$\mathcal{F}_{\mathbb{S}^2}$} (ads);        
    \draw [->] (ads) -- node[left] {$\mathcal{F}_{\mathbb{R}^2}$} (sas);
    \draw [->] (sds) -- node[] {$\mathcal{F}_{\mathbb{S}^2}$} (sas);
    \draw [->] (sad) -- node[left] {$\mathcal{F}_{\mathbb{R}^2}$} (sds);
    \draw [->] (is) -- node[left] {$\mathcal{F}^{-1}_{\mathbb{R}^2}$} (irr);

    \draw [->] (ads) -- node[text width=4cm, align=center] (cap1) {Angular dipole\\ transfer function\\ $\sum_{\ell m}\gray{H_\ell^m(\rd - \ro)}F_\ell^m(\ro)$} (irr);
    \draw [->] (sad.north) -- ++(0,1.0cm) -- node[right=2.cm, above, align=center] {Dipole point\\ spread function\\ $\int_{\mbb{S}^2}d\so\, \int_{\mbb{R}^2} d\ro\, \gray{h(\rd - \ro, \so)}f(\ro, \so)$} ++(13cm,0) --  (irr.north);
    \draw [->] (sds.south) -- ++(0,-1.0cm) -- node[right=1.95cm, above, align=center] {Spatial dipole\\ transfer function\\ $\int_{\mbb{S}^2}d\so\, \gray{H(\bs{\nu}, \so)}F(\bs{\nu}, \so)$} ++(13cm,0) -- (is.south);
    \draw [->] (sas) -- node[text width=4cm, align=center] {Spatio-angular dipole\\ transfer function\\ $\sum_{\ell m}\gray{\mathsf{H}_\ell^m(\bs{\nu})}\mathsf{F}_\ell^m(\bs{\nu})$} (is);

    % Dashed box
    \draw [dashed] (sad.north west) ++(-0.5cm,3cm) -- node[below, align=center] {Object space\\ $\mbb{U} = \mbb{L}_2(\mbb{R}^2\times\mbb{S}^2)$} ++(7.5cm,0) -- ++(0,-9cm) -- ++(-7.5cm,0) -- ++(0, 9cm);
    \draw [dashed] (irr.north west) ++(-0.5cm,3cm) -- node[below, align=center] {Data space\\ $\mbb{V} = \mbb{L}_2(\mbb{R}^2)$} ++(3.4cm,0) -- ++(0,-9cm) -- ++(-3.4cm,0) -- ++(0, 9cm);
      \end{tikzpicture}
\end{document}