### 2018-11-07 papers/basic-theory/figures/transfer-functions/transfer-functions.tex

\documentclass[crop,tikz]{standalone}% 'crop' is the default for v1.0, before it was 'preview'
% \usetikzlibrary{...}% tikz package already loaded by 'tikz' option
\usepackage{tikz}
\usepackage{amsmath, amsthm, amssymb, amsfonts}
\usetikzlibrary{shapes,arrows.meta}
\usepackage{newtxtext,newtxmath,courier}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}

% Talon's custom commands
\newcommand{\argmin}{\arg\!\min}
\newcommand{\me}{\mathrm{e}}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mf}[1]{\mathfrak{#1}}
\providecommand{\mc}[1]{\mathcal{#1}}
\providecommand{\ro}{\mathbf{\mathbf{r}}_o}
\providecommand{\so}{\mathbf{\hat{s}}_o}
\providecommand{\rp}{\mathbf{r}_p}
\providecommand{\rbm}[1]{r_b^{\text{m}}}
\providecommand{\rd}{\mathbf{r}_d}
\providecommand{\rdf}{\mathpzc{r}_d}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\mbb}[1]{\mathbb{#1}}
\providecommand{\bs}[1]{\boldsymbol{#1}}
\providecommand{\bv}{\bs{\nu}}
\providecommand{\bvp}{\bs{\nu}_p}
\providecommand{\bsh}[1]{\hat{\boldsymbol{#1}}}
\providecommand{\nan}{\left(\frac{\text{NA}}{n_o}\right)}
\providecommand{\lmsum}{\sum_{l=0}^\infty\sum_{m=-l}^{l}}
\providecommand{\intr}[1]{\int_{\mbb{R}^{#1}}}
\providecommand{\ints}[1]{\int_{\mbb{S}^{#1}}}
\DeclareFontFamily{OT1}{pzc}{}
\DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\newcommand{\eqname}[1]{\tag*{#1}}
\newcommand*\widefbox[1]{\fbox{\hspace{1em}#1\hspace{1em}}}

\tikzstyle{block} = [draw, fill=white, rectangle, minimum height=4em, minimum width=10em, align=center]
\begin{document}
  \begin{tikzpicture}[auto, node distance=2.5cm,>={Latex[length=2mm]}]
    \tikzset{font={\fontsize{9pt}{12}\selectfont}}
    \node [block, align=center] (sad) {$h(\rd - \ro, \so)$\\ Dipole point\\ spread function};
    \node [block,  right of=sad, node distance=5.5cm] (ads) {$H_\ell^m(\rd - \ro)$\\ Angular dipole\\ transfer function};
    \node [block,  left of=sad, node distance=5.5cm] (e) {$\mb{c}(\rd - \ro, \so)$\\ Coherent dipole \\ spread function};
      \node [block, below of=ads, node distance=2.5cm] (sas) {$\mathsf{H}_\ell^m(\bs{\nu})$\\ Spatio-angular\\ dipole transfer function};
      \node [block, below of=sad, node distance=2.5cm] (sds) {$H(\bs{\nu}, \so)$\\ Dipole\\ transfer function};
        \node [block,  left of=sds, node distance=5.5cm] (e2) {$\mb{C}(\bvp,\so)$\\ Coherent dipole \\transfer function};
      \draw [->] (sad) -- node[shift={(0,0)}] {$\mathcal{F}_{\mathbb{S}^2}$} (ads);
      \draw [->] (e) -- node[shift={(0,.75)}] {$|\mb{c}(\rd - \ro, \so)|^2$} (sad);
      % \draw [->] (e2) -- node[] {$\int_{\mbb{R}^2}d\rb \mb{E}_{\text{bfp}}(\rb, \so)\mb{E}_{\text{bfp}}^\dagger(\rb - \bs{\nu}, \so)$} (sds);
            \draw [->] (e2) -- node[below, shift={(0,-0.75)}] {$\int_{\mbb{R}^2}d\bvp\, \mb{C}(\bvp, \so)\mb{C}^\dagger(\bvp - \bs{\nu}, \so)$} (sds);        
      \draw [->] (ads) -- node[left, shift={(0,0)}] {$\mathcal{F}_{\mathbb{R}^2}$} (sas);
          \draw [->] (e) -- node[left] {$\mathcal{F}_{\mathbb{R}^2}$} (e2);
    \draw [->] (sds) -- node[below, shift={(0,0)}] {$\mathcal{F}_{\mathbb{S}^2}$} (sas);
    \draw [->] (sad) -- node[left] {$\mathcal{F}_{\mathbb{R}^2}$} (sds);
      \end{tikzpicture}
\end{document}