### 2018-11-21 papers/basic-theory/paper/paper.tex
New special functions, added simulation figures.

\documentclass[]{osa-article}

%% Select the journal you're submitting to
%% oe, boe, ome, osac, osajournal
\journal{osajournal}
% Key:
% Express journals must have the correct journal selected:
% {oe} Optics Express
% {boe} Biomedical Optics Express
% {ome} Optical Material Express
% {osac} OSAC Continuum
% Other OSA journals may use:
% {osajournal} Applied Optics, Advances in Optics and Photonics, Journal of the Optical Society of America A/B, Optics Letters, Optica, Photonics Research

% Uncomment if submitting to Photonics Research.
% ONLY APPLICABLE FOR \journal{osajournal}
% \setprjcopyright

% Set the article type
\articletype{Research Article}
% Note that article type is not required for Express journals (OE, BOE, OME and OSAC)

% Talon's custom libraries
\usepackage{empheq}
\usepackage{bm}

% Talon's custom commands
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\argmin}{\arg\!\min}
\newcommand{\me}{\mathrm{e}}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}} 
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\msf}[1]{\mathsf{#1}}
\providecommand{\mf}[1]{\mathfrak{#1}}
\providecommand{\mc}[1]{\mathcal{#1}}
\providecommand{\ro}{\mathbf{\mathbf{r}}_o}
\providecommand{\so}{\mathbf{\hat{s}}_o}
\providecommand{\rp}{\mathbf{r}_p}
\providecommand{\rbm}[1]{r_b^{\text{m}}}
\providecommand{\rd}{\mathbf{r}_d}
\providecommand{\rdf}{\mathpzc{r}_d}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\mbb}[1]{\mathbb{#1}}
\providecommand{\bs}[1]{\boldsymbol{#1}}
\providecommand{\bv}{\bs{\nu}}
\providecommand{\taup}{\bs{\tau}}
\providecommand{\bsh}[1]{\hat{\boldsymbol{#1}}}
\providecommand{\nan}{\left(\frac{\text{NA}}{n_o}\right)}
\providecommand{\lmsum}{\sum_{\ell=0}^\infty\sum_{m=-\ell}^{\ell}}
\providecommand{\intr}[1]{\int_{\mbb{R}^{#1}}}
\providecommand{\ints}[1]{\int_{\mbb{S}^{#1}}}
\providecommand{\tb}[1]{\textcolor{blue}{#1}}
% \DeclareFontFamily{OT1}{pzc}{}
% \DeclareFontShape{OT1}{pzc}{m}{it}{<-> s * [1.10] pzcmi7t}{}
% \DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\newcommand{\eqname}[1]{\tag*{#1}}
\newcommand*\widefbox[1]{\fbox{\hspace{1em}#1\hspace{1em}}}

\begin{document}

\title{Spatio-angular fluorescence microscopy I. Basic theory (Draft: \today)}

\author{Talon Chandler,\authormark{1,*} Authors TBD,\authormark{2,*}Hari Shroff\authormark{2,*}, Rudolf Oldenbourg\authormark{2,*}, and Patrick La Rivi\`ere \authormark{1}}

\address{\authormark{1}University of Chicago, Department of Radiology, Chicago, Illinois 60637, USA\\
\authormark{2}Publications Department, The Optical Society, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA\\
\authormark{3}Currently with the Department of Electronic Journals, The Optical Society, 2010 Massachusetts Avenue NW, Washington, DC 20036, USA}

\email{\authormark{*}talonchandler@talonchandler.com} %% email address is required

%\homepage{talonchandler@talonchandler.com} %% author's URL, if desired

%%%%%%%%%%%%%%%%%%% abstract %%%%%%%%%%%%%%%%
%% [use \begin{abstract*}...\end{abstract*} if exempt from copyright]

\begin{abstract*}
  We introduce the basic elements of a spatio-angular theory of fluorescence
  microscopy. We start by modeling an aplanatic microscope imaging an ensemble
  of in-focus fluorescent dipoles as a linear Hilbert-space operator with domain
  $\mbb{L}_2(\mbb{R}^2\times\mbb{S}^2)$ and range $\mbb{L}_2(\mbb{R}^2)$, and we
  express the operator in terms of four different sets of object-space basis
  functions. We show that the operator takes a particularly convenient form when
  expressed in a basis of spatial and spherical harmonics---a form we call the
  dipole spatio-angular transfer function. We demonstrate our formalism by
  analyzing a single-view fluorescence microscope without using the monopole or
  scalar approximations. We show that this imaging system has an angular band
  limit, and we demonstrate the value of the transfer function approach by
  efficiently simulating the imaging process with simple phantoms. Notably, we
  show that information about the out-of-plane orientation of ensembles of
  in-focus fluorophores is recorded by fluorescence microscopes. We discuss the
  implications of our analysis for all quantitative fluorescence microscopy
  studies and lay out a path towards a complete theory.
\end{abstract*}

\section{Introduction}

Fluorescence microscopes are widely used in the biological sciences for
measuring the spatial distribution of fluorophores throughout a sample [cite].
While an unprocessed fluorescence micrograph reports the approximate
distribution of fluorophores throughout a sample, all microscopes are
diffraction limited, so the image is a blurred version of the true
fluorophore distribution.

Restoration techniques attempt to recover the true distribution of fluorophores
using the measured data and a model of the imaging process. A model of the
imaging process can be obtained theoretically (by mathematically modeling the
instrument under idealized conditions), experimentally (by measuring the
instrument's response to a known input), or by a combination of theory and
experiment (by measuring parameters of an instrument model). In all cases the
accuracy of the restored fluorophore distribution is limited by the accuracy of
the imaging model. All theoretical imaging models make simplifying
approximations that limit the accuracy of restorations, so it is important to
verify that the approximations introduce an acceptable level of error. This work
investigates the errors introduced by two common approximations in models of
fluorescence microscopes---the \textit{monopole approximation} and the
\textit{scalar approximation}.

The widely-used monopole model treats fluorophores as molecules that absorb and
emit light isotropically. Despite their use in models, monopole
absorbers/emitters do not exist in nature. All real fluorophores absorb and emit
light anisotropically, and in this work we investigate the more realistic dipole
model of fluorophores. Since the absorption and emission patterns of many widely
used fluorophores including green fluroescent protein (GFP) are known to be well
described by the dipole model [cite Inoue], we start by considering the dipole
model before higher-order multipole models.

The scalar approximation is another widely-used approximation in fluorescence
microscopy. The scalar approximation models the propagation of light using a
scalar-valued field $U$ instead of the more realistic vector-valued electric
field $\mb{E}$. Although the scalar model yields realistic results in many
cases, the scalar field ignores polarization-dependent effects. The absorption
and emission response of dipoles strongly depends on the polarization of the
fields, so we need to use an electromagnetic model of light propagation to
accurately model a microscope imaging dipoles.

This work lies at the intersection of three classes of fluorescence microscopy:
(1) ensemble imaging techniques where many fluorophores are imaged in parallel
and the goal is to find their concentration throughout the sample, (2) single
molecule localization microscopy (SMLM) techniques where fluorophores are sparse
in the sample and the goal is to find the position of each fluorophore, and (3)
polarized fluorescence techniques where many fluorophores are imaged in parallel
and the goal is to find their concentration and orientation. We briefly review
these three classes and focus on their use of the monopole and scalar
approximations.

The large majority of fluorescence microscopes are used to image ensembles of
fluorophores, and many widely used restoration techniques make either or both of
the monopole and scalar approximations. The Gibson-Lanni model \cite{gibson89}
and the Born and Wolf model \cite{born1980} are two examples of scalar models
that are widely used to restore data collected by fluorescence microscopes, and
the Richards and Wolf model \cite{richards} is another widely-used model that
accounts for vectorial diffraction. All three of these models describe how
optical fields propagate through an optical system, but they are agnostic to the
object that is being imaged. In other words, these models are \textit{optical}
models, and when they are used to model fluorescent objects they implicitly make
the monopole approximation.

The SMLM community has pioneered the use of rigorous electromagnetic models for
dipole imaging systems that do not adopt the scalar or monopole approximations
\cite{backer2014, lieb2004, nov2006}. When a single molecule fluoresces, its
irradiance pattern on the detector is strongly dependent on the orientation of
the emitting molecule. Backlund et al. \cite{backlund2014} have shown have shown
that ignoring the orientation of fluorophores can bias position estimates, so
the most accurate SMLM experiments must jointly estimate the position and
orientation of each molecule. Some SMLM experiments use polarizers to estimate
the orientation of single molecules more accurately [Backer DNA paper, other
Moerner papers], and these works do not adopt the scalar or monopole
approximations.

A smaller community of microscopists is interested in measuring the orientation
and position of ensembles of fluorophores \cite{mehta2016} [Forkey, Goldman,
Moerner, Oldenbourg]. These techniques typically use polarized illumination or
polarized detection to make multiple measurements of the same object. Current
angular restoration techniques use a model of the dipole excitation and emission
processes [Fourkas] to recover the orientation of fluorophores using pixel-wise
arithmetic, but these techniques do not perform any spatial restoration so they
do not use all of the available information.

In this work we place these three communities on a common theoretical footing.
We begin in Section \ref{sec:theory} by formulating the imaging problem
mathematically. We start with an abstract description of fluorescence imaging
systems and extend the usual monopole imaging model to dipoles. Along the way we
introduce transfer functions that will allow us to efficiently analyze
spatio-angular microscopes. In Section \ref{sec:results} we use the mathematical
tools we developed in Section \ref{sec:theory} to analyze and simulate a
single-view epi-fluorescence microscope, and we demonstrate our model with four
simple phantoms. Finally, in Section \ref{sec:discussion} we discuss the results
and their broader implications.

\section{Theory}\label{sec:theory}
We begin our analysis with the abstract Hilbert space formalism of Barrett and
Myers \cite{barrett2004}. Our first task is to formulate the imaging process as
a mapping between two Hilbert spaces $\mc{H}: \mbb{U} \rightarrow \mbb{V}$, where
$\mbb{U}$ is a set than contains all possible objects, $\mbb{V}$ is a set that
contains all possible datasets, and $\mc{H}$ is a model of the instrument that
maps between these two spaces. We denote (possibly infinite-dimensional)
Hilbert-space vectors in $\mbb{U}$ with $\mb{f}$, Hilbert-space vectors in
$\mbb{V}$ with $\mb{g}$, and the mapping between the spaces with
\begin{align}
  \mb{g} = \mc{H}\mb{f}.
\end{align}

Once we have identified the spaces $\mbb{U}$ and $\mbb{V}$, we can start
expressing the mapping between the spaces in a specific object-space and
data-space basis. In most cases the easiest mapping to find uses a
delta-function basis---we expand object and data space into delta functions,
then express the mapping as an integral transform. After finding this mapping we
can start to investigate the same mapping in different bases.

The above discussion is quite abstract, but it is a powerful point of view that
will enable us to unify the analysis of spatio-angular fluorescence microscopes.
In Section \ref{sec:monopole} we will demonstrate the formalism by examining a
familiar monopole imaging model, and we will demonstrate the mapping between
object and data space in two different bases. In Section \ref{sec:dipole} we
will extend the monopole imaging model to dipoles and examine the mapping in
four different bases.

\subsection{Monopole imaging in different bases}\label{sec:monopole}
We start by considering a microscope that images a field of in-focus monopoles
by recording the irradiance on a two-dimensional detector. This section treads
familiar ground, but it serves to establish the concepts and notation that will
be necessary when we extend to the dipole case.

We can represent the object as a function that assigns a real number to each
point on a plane, so we identify object space as
$\mbb{U} = \mbb{L}_2(\mbb{R}^2)$---the set of square-integrable functions on the
two-dimensional plane. Similarly, we have a two-dimensional detector that
measures a real number at each point on a plane, so data space is the same set
$\mbb{V} = \mbb{L}_2(\mbb{R}^2)$.

Next, we name the representations of our object and data in a specific basis. In
a delta function basis the object can be represented by a function $f(\ro)$
called the \textit{monopole density}---the number of monopoles per unit area at
the two-dimensional position $\ro$. Similarly, in a delta function basis the
data can be represented by a function $g'(\rd')$ called the
\textit{irradiance}---the power received by a surface per unit area at position
$\rd'$. It may seem pedantic to emphasize that these representations are in a
delta function basis, but we highlight this point to make it clear that the
delta function basis is not special---we can choose a different basis at will.

A reasonable starting point is to assume that the relationship between the
object and the data is \textit{linear}---this is usually true in fluorescence
microscopes since fluorophores emit incoherently, so a scaled sum of
fluorophores will result in a scaled sum of the irradiance patterns created by
the individual fluorophores. If the mapping is linear, we can write the
irradiance as a weighted integral over a field of monopoles
\begin{align}
g'(\rd') = \int_{\mbb{R}^2}d\ro\, h'(\rd',\ro)f(\ro), \label{eq:fwdmono}
\end{align}
where $h'(\rd{}', \ro{})$ is the irradiance at position $\rd'$ created by a
point source at $\ro$.

Next, we assume that the optical system is \textit{aplanatic}---Abbe's sine
condition is satisfied and on-axis points are imaged without aberration. Abbe's
sine condition guarantees that off-axis points are imaged without spherical
aberration or coma \cite{mansuripur2009}, so the imaging system can be modeled
within the field-of-view of the optical system as a magnifier with
shift-invariant blur
\begin{align}
  g'(\rd') = \int_{\mbb{R}^2}d\ro\, h'(\rd' - m\ro)f(\ro), \label{eq:nonconv}
\end{align}
where $m$ is a magnification factor.

We can simplify our analysis by changing coordinates and writing Eq.
\ref{eq:nonconv} as a convolution \cite{barrett2004}. We define a demagnified
detector coordinate $\rd = \rd'/m$ and a normalization factor that corresponds
to the total power incident on the detector plane due to a point source
$P_{\text{mono}} = \int_{\mbb{R}^2}d\mb{r}\,h'(m\mb{r})$ where
$\mb{r} = \rd - \ro$. We use these scaling factors to define the
\textit{monopole point spread function} as
\begin{align}
  h(\rd - \ro) = \frac{h'(m[\rd - \ro])}{P_{\text{mono}}},
\end{align}
and the \textit{scaled irradiance} as
\begin{align}
  g(\rd) = \frac{g'(m\rd)}{P_{\text{mono}}}.
\end{align}
With these definitions we can express the mapping between the object and the
data as a familiar convolution
\begin{align}
  g(\rd) = \int_{\mbb{R}^2}d\ro\, h(\rd - \ro)f(\ro).  \label{eq:lsi}
\end{align}
We will mostly be using scaled coordinates, so we have adopted a slightly
unusual convention of using primes to denote unscaled coordinates.

We have chosen to normalize the monopole point spread function so that
\begin{align}
  \int_{\mbb{R}^2}d\mb{r}\, h(\mb{r}) = 1. \label{eq:norm}
\end{align}
The monopole point spread function corresponds to a measurable irradiance, so it
is always real and positive.

\begin{figure}
  \centering
  \includegraphics[scale=1.0]{../figures/monopole-block/monopole-block.pdf}
  \caption{The mapping between the object and data space of a monopole
    fluorescence microscope can be computed in two different bases---a delta
    function basis and a spatial harmonic basis. The change of basis can be
    computed with a two-dimensional Fourier transform denoted
    $\mathcal{F}_{\mbb{R}^2}$.}
     \label{fig:monopole-block}      
\end{figure}

The mapping between the object and the data in a linear shift-invariant imaging
system takes a particularly simple form in a spatial harmonic basis. If we apply
the Fourier convolution theorem to Eq. \ref{eq:lsi} we find that
\begin{align}
  G(\bv) = H(\bv)F(\bv),\label{eq:freq}
\end{align}
where we define the \textit{scaled irradiance spectrum} as
\begin{align}
  G(\bv) = \int_{\mbb{R}^2}d\mb{r}\, g(\mb{r})\, \text{exp}(-2\pi i\mb{r}\cdot\bv),
\end{align}
the \textit{monopole transfer function} as
\begin{align}
  H(\bv) = \int_{\mbb{R}^2}d\mb{r}\, h(\mb{r})\, \text{exp}(-2\pi i\mb{r}\cdot\bv),\label{eq:otf}
\end{align}
and the \textit{monopole spectrum} as
\begin{align}
    F(\bv) = \int_{\mbb{R}^2}d\mb{r}\, f(\mb{r})\, \text{exp}(-2\pi i\mb{r}\cdot\bv).
\end{align}
The monopole point spread function is normalized and real, so we know
that the monopole transfer function is normalized $H(0) = 1$ and
conjugate symmetric $H(-\bv) = H^*(\bv)$ where $z^*$ denotes the complex
conjugate of $z$.

Notice that Eqs. \ref{eq:lsi} and \ref{eq:freq} are expressions of the same
mapping between object and data space in different bases. Figure
\ref{fig:monopole-block} summarizes the relationship between object and data
space in both bases.

We have been careful to use the term \textit{monopole transfer function} instead
of the commonly-used term \textit{optical transfer function}. We reserve the
term \textit{optical transfer function} for optical systems---the optical
transfer function maps between an input irradiance spectrum and an output
irradiance spectrum in an optical system. We can use optical transfer functions
to model the propagation of light through a microscope, but ultimately we are
always interested in the object, not the light emitted by the object. We will
find the distinction between the optical transfer function and the object
transfer function to be especially valuable when we consider dipoles later in
this section.

\subsubsection{Monopole coherent transfer functions}
Although the Fourier transform can be used to calculate the monopole transfer
function directly from the monopole point spread function, there is a well-known
alternative that exploits coherent transfer functions. The key idea is that the
monopole point spread function can always be written as the absolute square of a
scalar-valued \textit{monopole coherent spread function} defined as
\begin{align}
  |c(\rd - \ro)|^2 = h(\rd - \ro). \label{eq:absquarescalar}
\end{align}
Physically, the monopole coherent spread function corresponds to the
scalar-valued field on the detector with appropriate scaling.

Since the monopole transfer function is given by the two-dimensional Fourier
transform of the monopole point spread function
\begin{align}
  H(\bv) = \int_{\mbb{R}^2}d\mb{r}\,h(\mb{r})\,\text{exp}[-2\pi i\mb{r}\cdot\bv],
\end{align}
we can plug in Eq. \ref{eq:absquarescalar} and use the autocorrelation theorem to
rewrite the monopole transfer function as
\begin{align}
  H(\bv) = \int_{\mbb{R}^2}d\taup\, C(\taup)C^*(\taup - \bv), 
\end{align}
where we have introduced the
\textit{monopole coherent transfer function} as the two-dimensional Fourier
transform of the monopole coherent spread function
\begin{align}
  C(\taup) = \int_{\mbb{R}^2}d\mb{r}\, c(\mb{r})\,\text{exp}[-2\pi i\mb{r}\cdot\taup].
\end{align}
Physically, the monopole coherent transfer function corresponds to the
scalar-valued field in a Fourier plane of the detector with appropriate scaling.
% We have chosen the symbol $\taup$ as the coordinate for the coherent
% transfer function because the Fourier plane of the detector is often called the
% \textit{pupil plane}.

The coherent transfer function provides a valuable shortcut for analyzing
microscopes since it is often straightforward to calculate the field in a
Fourier plane of the detector. A typical approach for calculating the transfer
functions is to (1) calculate the field in a Fourier plane of the detector, (2)
scale the field to find the monopole coherent transfer function, then (3) use
the relationships in Figure \ref{fig:monopole-transfer-functions} to calculate
the other transfer functions.

\begin{figure}
  \centering
  \includegraphics[scale=1.0]{../figures/monopole-transfer-functions/monopole-transfer-functions.pdf}
  \caption{The incoherent monopole transfer functions are related by a
    two-dimensional Fourier transform (right column). The coherent monopole
    transfer functions (left column) can be used to simplify the calculation of
    the remaining transfer functions.}
   \label{fig:monopole-transfer-functions}
 \end{figure}

\subsection{Dipole imaging in different bases}\label{sec:dipole}
Now we consider a microscope imaging a field of in-focus dipoles by recording
the irradiance on a two-dimensional detector. A function that assigns a real
number to each point on a plane is not sufficient to specify a field of dipoles
because the dipoles can have different orientations. To represent the object we
need to extend object space to $\mbb{U} = \mbb{L}_2(\mbb{R}^2\times\mbb{S}^2)$
where $\mbb{S}^2$ is the two-dimensional sphere (the usual sphere embedded in
$\mbb{R}^3$).

In a delta function basis the object can be represented by a function
$f(\ro, \so)$ called the \textit{dipole density}---the number of dipoles at
position $\ro{}$ per unit area oriented along $\so{}$ per unit solid angle.
Similar to the monopole case, we model the mapping between the object and the
irradiance in a delta function basis as an integral transform
\begin{align}
  g'(\rd') = \int_{\mbb{S}^2}d\so\int_{\mbb{R}^2}d\ro\, h'(\rd', \ro, \so)f(\ro, \so). 
\end{align}
where $h'(\rd', \ro, \so)$ is the irradiance at position $\rd'$ created by a
point source at $\ro$ with orientation $\so$. If the optical system is aplanatic
we can write the integral transform as
\begin{align}
  g'(\rd') = \int_{\mbb{S}^2}d\so\int_{\mbb{R}^2}d\ro\, h'(\rd' - m\ro, \so)f(\ro, \so). 
\end{align}

We define the same demagnified detector coordinate $\rd = \rd'/m$ and a new
normalization factor that corresponds to the total power incident on the
detector due to a spatial point source with an angularly uniform distribution of
dipoles
$P_\text{dip} = \int_{\mbb{S}^2}d\mh{s}\int_{\mbb{R}^2}d\mb{r}\, h'(m\mb{r},
\so)$. We use these scaling factors to define the
\textit{dipole point spread function} as
\begin{align}
  h(\rd - \ro, \so) = \frac{h'(m[\rd - \ro], \so)}{P_\text{dip}}, 
\end{align}
and the \textit{scaled irradiance} as 
\begin{align}
  g(\rd) = \frac{g'(m\rd)}{P_\text{dip}}. 
\end{align}
With these definitions we can express the mapping between the object and the
data as
\begin{align}
g(\rd{}) = \int_{\mbb{S}^2}d\so{}\int_{\mbb{R}^2}d\ro{}\, h(\rd{} -\ro{}, \so{})f(\ro, \so). \label{eq:odpsf}
\end{align}
Eq. \ref{eq:odpsf} is a key result because it represents the mapping between
object space and data space in a delta function basis. The integrals in
\ref{eq:odpsf} would be extremely expensive to compute for an arbitrary object,
but the integrals simplify to an efficient sum if the object consists of
spatially fixed dipoles in a fixed orientation. In other words, Eq.
\ref{eq:odpsf} is ideal for simulating and analyzing objects that are spatially
and angularly sparse; e.g. single fluorophores that are rigidly attached to a
structure .

Similar to the monopole case, we have chosen to normalize the dipole point
spread function so that
\begin{align}
  \int_{\mbb{S}^2}d\so\int_{\mbb{R}^2}d\mb{r}\, h(\mb{r}, \so) = 1. 
\end{align}
The dipole point spread function is a measurable quantity, so it is real
and positive.

\subsubsection{Spatial dipole transfer function}
We can make our first change of basis by applying the Fourier-convolution
theorem to Eq. \ref{eq:odpsf} which yields
\begin{align}
G(\bv) = \int_{\mbb{S}^2}d\so\, H(\bv, \so)F(\bv, \so) \label{eq:odotf},
\end{align}
where we define the \textit{spatial dipole transfer function} as
  \begin{align}
  H(\bv, \so) &= \int_{\mbb{R}^2}d\mb{r}\, h(\mb{r}, \so)\, \text{exp}(-2\pi i\mb{r}\cdot\bv),
  \end{align}
  and the \textit{spatial dipole spectrum} as
  \begin{align}
  F(\bv, \so) &= \int_{\mbb{R}^2}d\mb{r}\, f(\mb{r}, \so)\, \text{exp}(-2\pi i\mb{r}\cdot\bv). 
  \end{align}
  Since the dipole point spread function is normalized and real, we know that
  the spatial dipole transfer function is normalized
  $\int_{\mbb{S}^2}d\mh{s}\, H(0, \so) = 1$ and conjugate symmetric
  $H(-\bv, \so) = H^*(\bv, \so)$.

  This basis is ideal for simulating and analyzing objects that are angularly
  sparse and spatially dense; e.g. rod structures that contain fluorophores in a
  fixed orientation.

  \begin{figure}
  \hspace{-5em}
  \includegraphics[scale=1.0]{../figures/dipole-block/dipole-block.pdf}
  \caption{The mapping between the object and data space of a dipole imaging
    system can be computed in four different bases---a delta function basis, a
    spatial-harmonic/angular-delta basis, a spatial-delta/spherical-harmonic
    basis, and a spatio-angular harmonic basis. The changes of basis can be
    computed with the two-dimensional Fourier transform denoted
    $\mathcal{F}_{\mbb{R}^2}$, and the spherical Fourier transform denoted
    $\mathcal{F}_{\mbb{S}^2}$.}
   \label{fig:dipole-block}      
    \end{figure}

\subsubsection{Dipole angular transfer function}
The spherical harmonics are another set of convenient basis functions that are
play the same role as complex exponentials in spatial transfer functions---see
Appendix \ref{sec:sph} for an introduction to the spherical harmonics. We can
change basis from spherical delta functions to spherical harmonics by applying
the generalized Plancharel theorem for spherical functions
\begin{align}
  \ints{2}d\mh{s}\, p(\mh{s})q^*(\mh{s}) = \lmsum P_\ell^m Q_\ell^{m*}, \label{eq:plan}
\end{align}
where $p(\mh{s})$ and $q(\mh{s})$ are arbitrary functions on the sphere,
$P_\ell^m$ and $Q_\ell^m$ are their spherical Fourier transforms defined by
\begin{align}
  P_\ell^m = \int_{\mbb{S}^2}d\mh{s}\, p(\mh{s})Y_\ell^{m*}(\mh{s}),
\end{align}
and $Y_{\ell}^m(\mh{s})$ are the spherical harmonic functions defined in
Appendix \ref{sec:sph}. Eq. \ref{eq:plan} expresses the fact that scalar
products are invariant under a change of basis (see Eq. 3.78 of
\cite{barrett2004}). The left-hand side of Eq. \ref{eq:plan} is the scalar
product of $\mbb{L}_2(\mbb{S}^2)$ functions in a delta function basis and the
right-hand side is the scalar product of $\mbb{L}_2(\mbb{S}^2)$ functions in a
spherical harmonic function basis. Applying Eq. \ref{eq:plan} to Eq.
\ref{eq:odpsf} yields
\begin{align}
  g(\rd) = \lmsum H_\ell^m(\rd - \ro)F_\ell^m(\ro), \label{eq:atf-form}
\end{align}
where we have defined the \textit{dipole angular transfer function} as
\begin{align}
  H_\ell^m(\rd - \ro) = \int_{\mbb{S}^2}d\so\, h(\rd - \ro, \so)Y_{\ell}^{m*}(\so),\label{eq:atf-prep} 
\end{align}
and the \textit{dipole angular spectrum} as
\begin{align}
  F_\ell^m(\ro) = \int_{\mbb{S}^2}d\so\, f(\ro, \so)Y_{\ell}^{m*}(\so).
\end{align}
Since the dipole point spread function is normalized and real, we know
that the dipole angular transfer function is normalized
$\int_{\mbb{R}^2}d\mb{r}\, H_0^0(\mb{r}) = 1$ and conjugate symmetric
$H_\ell^{-m}(\mb{r}) = (-1)^mH_\ell^{m*}(\mb{r})$.

This basis is well suited for simulating and analyzing objects that are
spatially sparse and angularly dense; e.g. rotating single fluorophores that are
spatially fixed, or many fluorophores that are close together with varying
orientations.

\subsubsection{Spatio-angular dipole transfer function}
We can arrive at our final basis in two ways: by applying the generalized
Plancharel theorem for spherical functions to Eq. \ref{eq:odotf} or by applying
the Fourier convolution theorem to Eq. \ref{eq:atf-form}. We follow the
second path and find that
\begin{align}
G(\bv) = \lmsum \msf{H}_\ell^m(\bv)\msf{F}_\ell^m(\bv) \label{eq:saft},
\end{align}
where we have defined the \textit{dipole spatio-angular transfer function} as
  \begin{align}
  \msf{H}_\ell^m(\bv) &= \int_{\mbb{S}^2}d\so\, H(\bv, \so)Y_\ell^{m*}(\so),
  \end{align}
  and the \textit{dipole spatio-angular spectrum} as
  \begin{align}
  \msf{F}_\ell^m(\bv) &= \int_{\mbb{S}^2}d\so\, F(\bv, \so)Y_\ell^{m*}(\so).
  \end{align}
  Since the dipole point spread function is normalized and real, we know
  that the dipole spatio-angular transfer function is normalized
  $\msf{H}_0^0(0) = 1$ and conjugate symmetric
  $\msf{H}_\ell^{-m}(-\bv) = (-1)^m\msf{H}_\ell^{m*}(\bv)$.

  This basis is well suited for simulating and analyzing arbitrary samples
  because it exploits the band limit of the imaging system. We will see this
  advantage explicitly in Section \ref{sec:results} when we calculate a specific
  dipole spatio-angular transfer function and see that it has spatial and
  angular band limits.

  Figure \ref{fig:dipole-block} summarizes the relationship between the four
  bases that we can use to compute the image of a field of dipoles. We reiterate
  that all four bases may be useful depending on the sample.

\subsubsection{Dipole coherent transfer functions}
Similar to the monopole case, there is efficient way to calculate the transfer
functions using coherent transfer functions. The dipole point spread function
can always be written as the absolute square of a vector-valued function called
the \textit{dipole coherent spread function}
\begin{align}
  |\mb{c}(\rd - \ro, \so)|^2 = h(\rd - \ro, \so). \label{eq:absquare2}
\end{align}
Physically, the dipole coherent spread function corresponds to the vector-valued
electric field on the detector with appropriate scaling. We need a vector-valued
coherent transfer function since the polarization of the field plays a
significant role in dipole imaging, so the point spread function cannot be
written as an absolute square of a scalar-valued function.

Since the dipole spatial transfer function is the two-dimensional
Fourier transform of the dipole point spread function
\begin{align}
  H(\bv) = \int_{\mbb{R}^2}d\mb{r}\,h(\mb{r}, \so)\,\text{exp}[-2\pi i\mb{r}\cdot\bv],
\end{align}
we can plug in Eq. \ref{eq:absquare2} and use the autocorrelation theorem to
rewrite the spatial dipole transfer function as 
\begin{align}
  H(\bv) = \int_{\mbb{R}^2}d\taup\,\mb{C}(\taup, \so)\mb{C}^\dagger(\taup - \bv, \so), 
\end{align}
where we have introduced the \textit{dipole coherent transfer function}
$\mb{C}(\taup, \so)$ as the two-dimensional Fourier transform of the dipole
coherent spread function
\begin{align}
  \mb{C}(\taup, \so) = \int_{\mbb{R}^2}d\mb{r}\, \mb{c}(\mb{r}, \so)\,\text{exp}[-2\pi i\mb{r}\cdot\taup].
\end{align}
Physically, the dipole coherent transfer function corresponds to the
vector-valued electric field created by a dipole oriented along $\so$ in a
Fourier plane of the detector with appropriate scaling. Similar to the monopole
case, we can calculate the orientation-dependent fields in a Fourier plane of
the detector, scale appropriately to find the dipole coherent transfer function,
then use the relationships in Figure \ref{fig:transfer-functions} to calculate
the other transfer functions.

\begin{figure}
  \hspace{-2em}
  \includegraphics[scale=1.0]{../figures/transfer-functions/transfer-functions.pdf}
  \caption{There is one incoherent transfer function for each set of
    object-space basis functions, and these transfer functions are related by
    two-dimensional and spherical Fourier transforms---see center and right
    columns. There are an additional pair of coherent transfer functions that
    are useful for calculating the incoherent transfer functions---see left column.}
   \label{fig:transfer-functions}
 \end{figure}

\section{Results}\label{sec:results}
In this section we begin to apply the tools we developed in the previous section
to analyze a specific fluorescence microscope. During our initial discussion we
only required the optical system to be aplanatic so that we could model it as a
shift-invariant magnifier. Now we will be more specific and consider an
aplanatic optical system in a \textit{$\mathit{4}f$ configuration} with an arbitrary
first lens (the objective lens) and a \textit{paraxial second lens} (the tube
lens) as shown in Figure \ref{fig:schematic}. A lens can be considered paraxial
if the angle $\alpha$ between the optical axis of the lens and the marginal ray
is small enough that $\sin\alpha \approx \alpha$. As a rule of thumb,
non-paraxial effects only become significant when the numerical aperture of a
lens exceeds 0.7 \cite{gu2000}, but this is only a rough guideline.

We start by defining \textit{pupil functions} for both monopole and dipole
imaging systems, and we explicitly relate the pupil functions to the previously
defined coherent transfer functions. After establishing the link between
physical calculations and the transfer functions, we calculate the monopole and
dipole transfer functions for imaging systems with a paraxial objective, and we
demonstrate our approach by simulating simple phantoms.

\begin{figure}[h]
 \centering
   \centering
   \includegraphics[width = 0.9\textwidth]{../figures/abstract-micro/abstract-micro.pdf}
   \caption{Schematic of an aplanatic imaging system in a $4f$ geometry with a
     paraxial tube lens. Since we have an aplanatic optical system we only need
     to consider the image created by on-axis objects.}
   \label{fig:schematic}
 \end{figure}

\subsection{Monopole pupil functions}
We define the \textit{monopole pupil function} $p(\rp)$ of the imaging system as
the field immediately following the pupil plane created by an on-axis
monopole where $\rp$ is an unscaled two-dimensional coordinate in the pupil
plane. In this section we will relate the monopole pupil function to the
transfer function we defined in Section \ref{sec:theory} by following a near
variant of the derivation in Barrett and Myers Section 9.7 \cite{barrett2004}.

Since monopoles emit scalar fields, the monpole pupil function is a
scalar-valued function. The optical system is aplanatic, so we can write the
field created at a point in the aperture $\rp$ by a monopole at position $\ro$
as
\begin{align}
   U_p(\rp, \ro) \propto p(\rp)\,\text{exp}\left[-2\pi i \frac{n_0}{\lambda f_0} \rp\cdot\ro \right]. \label{eq:pupil}
\end{align}
Eq. \ref{eq:pupil} is a restatement of the aplanatic condition for a $4f$
optical system---the field in the aperture plane can be written as the pupil
function multiplied by a linear phase factor that encodes the position of the
object.

Since the second lens is paraxial, we can model the relationship between the
field in the aperture plane and the field on the detector with a scaled Fourier
transform \cite{goodman1996}
\begin{align}
   U_d(\rd', \ro) \propto \int_{\mbb{R}^2}d\rp\, p(\rp)\,\text{exp}\left[-2\pi i \frac{n_0}{\lambda f_0} \rp\cdot\ro \right]\text{exp}\left[-2\pi i \frac{n_1}{\lambda f_1} \rp\cdot\rd' \right]. \label{eq:detectorfourier}
\end{align}
If we define $P(\taup)$ as the two-dimensional Fourier transform of the pupil function
then we can rewrite Eq. \ref{eq:detectorfourier} as
\begin{align}
  U_d(\rd', \ro) \propto P\left(\frac{n_0}{\lambda f_0}\ro + \frac{n_1}{\lambda f_1}\rd'\right),
\end{align}
which we can simplify further by writing in terms of
$m = -\frac{f_1n_0}{f_0n_1}$:
\begin{align}
  U_d(\rd' - m\ro) \propto P\left(\frac{n_1}{\lambda f_1}[\rd' - m\ro]\right).
\end{align}
The irradiance on the detector is the absolute square of the field so
\begin{align}
  h'(\rd' - m\ro) \propto \left|P\left(\frac{n_1}{\lambda f_1}[\rd' - m\ro]\right)\right|^2.
\end{align}
If we demagnify the coordinates with $\rd = \rd'/m$ and demagnify the irradiance
with $h(\rd - \ro) \propto h'(m[\rd - \ro])$ we find that the monopole point spread
function is related to the Fourier transform of the monopole pupil function by
\begin{align}
  h(\rd - \ro) \propto \left|P\left(-\frac{n_o}{\lambda f_o}[\rd - \ro]\right)\right|^2.
\end{align}
The monopole coherent spread function is the absolute square of the monopole point spread function so
\begin{align}
  c(\rd - \ro) \propto P\left(-\frac{n_o}{\lambda f_o}[\rd - \ro]\right).
\end{align}
Finally, the monopole coherent transfer function is the Fourier transform of the
monopole coherent spread function so
\begin{align}
  C(\taup) \propto p\left(\frac{\lambda f_o}{n_o}\taup\right). \label{eq:monopupil0}
\end{align}
Equation \ref{eq:monopupil0} is the key result of this section---the monopole
coherent transfer function is a scaled monopole pupil function.

\subsection{Dipole pupil function}
We define the \textit{dipole pupil function} $\mb{p}(\rp, \so)$ of the imaging
system as the electric field immediately following the pupil plane created by
an on-axis dipole oriented along $\so$. Since dipoles emit vector-valued
electric fields, the dipole pupil function is a vector-valued function. Almost
all of the arguments in the previous section carry over to the dipole case.
Briefly, we can write the electric field created at a point in the pupil
$\rp$ by a dipole at $\ro$ oriented along $\so$ as
 \begin{align}
   \mb{E}(\rp, \ro, \so) = \mb{p}(\rp, \so)\,\text{exp}\left[-2\pi i \frac{n_0}{\lambda f_0} \rp\cdot\ro \right]. \label{eq:pupil2}
 \end{align}
 The second lens is paraxial, so we can find the field on the detector with a
 Fourier transform
 \begin{align}
   \mb{E}_d(\rd', \ro, \so) \propto \int_{\mbb{R}^2}d\rp\, \mb{p}(\rp, \so)\,\text{exp}\left[-2\pi i \frac{n_0}{\lambda f_0} \rp\cdot\ro \right]\text{exp}\left[-2\pi i \frac{n_1}{\lambda f_1} \rp\cdot\rd' \right]. \label{eq:detectorfourier2}
\end{align}
Note that the Fourier transform of a vector field is the Fourier transform of
its scalar-valued orthogonal components, so Eq. \ref{eq:detectorfourier2}
specifies two two-dimensional Fourier transforms. We follow the same
manipulations as the previous section and find that the dipole coherent transfer
function is a scaled dipole pupil function
\begin{align}
  \mb{C}(\taup, \so) \propto \mb{p}\left(\frac{\lambda f_o}{n_o}\taup, \so\right). \label{eq:monopupildip}
\end{align}

\subsection{Special functions}
We adopt and generalize Bracewell's notation \cite{bracewell2004} for several
special functions which will simplify our calculations. First, we define a
\textit{rectangle function} as
\begin{align}
  \Pi(x) = 
  \begin{cases}
    1\quad \text{if}\quad |x| < \frac{1}{2},\\
    0\quad \text{else}.
  \end{cases}
\end{align}
We also define the \textit{$n^{th}$-order jinc function} as
\begin{align}
  \text{jinc}_n(r) = \frac{J_{n+1}(\pi r)}{2r},
\end{align}
where $J_{n+1}(r)$ is the $(n+1)^{th}$-order Bessel function of the first kind.

In Appendix \ref{sec:special} we derive the following two-dimensional Fourier
transform relationships between the jinc functions and the weighted rectangle
functions
\begin{align}
  i^{n}\left\{\substack{
  \text{exp}(in\phi_r)\\
    \cos(n\phi_r)\\
    \sin(n\phi_r)
  }\right\}\text{jinc}_n(r) &\stackrel{\mathcal{F}_{\mbb{R}^2}}{\longrightarrow} (2\nu)^{n}\left\{\substack{
    \text{exp}(in\phi_{\nu})\\
    \cos(n\phi_{\nu})\\
    \sin(n\phi_{\nu})
  }\right\}\Pi(\nu), \label{eq:jincrect2}
  \end{align}
where the entries inside the curly braces are to be taken one at a time.

Finally, we define the \textit{$n^{th}$-order chat function} as the
two-dimensional Fourier transform of the squared $n^{th}$-order jinc function
\begin{align}
  \text{jinc}_n^2(r) \stackrel{\mathcal{F}_{\mbb{R}^2}}{\longrightarrow} \text{chat}_n(\nu).
\end{align}

In Appendix \ref{sec:special} we show that the zero and first order chat
functions can be written in closed form as
\begin{align}
  \text{chat}_0(x) &=  \frac{1}{2}\left[\cos^{-1}|x| - |x|\sqrt{1 - x^2}\right]\Pi\left(\frac{x}{2}\right),\\
  \text{chat}_1(x) &=  \frac{1}{2}\left[\cos^{-1}|x| - |x|(3-2x^2)\sqrt{1 - x^2}\right]\Pi\left(\frac{x}{2}\right).  
\end{align}

\subsection{Monopole transfer functions}
Our first step towards the monopole transfer functions is to calculate the
monopole pupil function and coherent transfer function. Several works
\cite{petrov2017, backlund2018} have modeled an aplanatic fluorescence
microscope imaging monopole emitters with the scalar pupil function
\begin{align}
  p(\rp) \propto \tilde{C}\left(\frac{r_p}{f_o}\right)\Pi\left(\frac{r_p}{2f_o\sin\alpha}\right), 
\end{align}
where
\begin{align}
  \tilde{C}(x) = (1 - x^2)^{-1/4} = 1 + \frac{x^2}{4} + \frac{5x^4}{32} + \cdots. 
\end{align}
The $\tilde{C}(x)$ function models the radial dependence of the field and
ensures that power is conserved on either side of an aplanatic objective, and
the rectangle function models the aperture stop of the objective. Applying Eq.
\ref{eq:monopupil0} and collecting constants we find that the coherent monopole
transfer function is
\begin{align}
  C(\taup) \propto \tilde{C}\left(\frac{2\text{NA}}{n_o}\frac{\tau}{\nu_c}\right)\Pi\left(\frac{\tau}{\nu_c}\right), \label{eq:coherentmonopole}
\end{align}
where $\text{NA} = n_o\sin\alpha$ and $\nu_c = 2\text{NA}/\lambda$. This
coherent transfer function models objectives with an arbitrary numerical
aperture, but for our initial analysis we restrict ourselves to the paraxial
regime. We retain only zero and first order radial terms to find that  
\begin{align}
  C(\taup) \stackrel{(p)}{\propto} \Pi\left(\frac{\tau}{\nu_c}\right), 
\end{align}
where $(p)$ indicates that we have used the paraxial approximation for the
objective lens.

We can find the monopole coherent spread function by taking the inverse Fourier
transform of the monopole coherent transfer function
\begin{align}
  c(\mb{r}) \stackrel{(p)}{\propto} \text{jinc}_0(\nu_c r).
\end{align}

The monopole point spread function is the (normalized) absolute square of the
monopole coherent spread function so,
\begin{align}
  h(\mb{r}) \stackrel{(p)}{=} \frac{4}{\pi}\text{jinc}_0^2(\nu_c r)
\end{align}
which is the well-known \textit{Airy disk}.

Finally, we can calculate the monopole transfer function as the two-dimensional Fourier transform of the monopole point spread function (or the autocorrelation
of the coherent transfer function) and find that 
\begin{align}
  H(\bs{\nu}) \stackrel{(p)}{=} \frac{4}{\pi}\text{chat}_0\left(\frac{\nu}{\nu_c}\right).
\end{align}

 \subsection{Dipole transfer functions}
 To calculate the dipole transfer function we proceed similarly to the monopole
 case---we find the pupil function, scale to find the coherent dipole transfer
 function, then calculate the remaining transfer functions.

 Backer and Moerner \cite{backer2014} have calculated the dipole pupil function
 for a high-NA objective as
 \begin{align}
   \mb{p}(\rp, \so) \propto
   \begin{bmatrix}
     \tilde{C}_0\left(\frac{r_p}{f_o}\right) + \tilde{C}_2\left(\frac{r_p}{f_o}\right)\cos(2\phi_p)&\tilde{C}_2\left(\frac{r_p}{f_o}\right)\sin(2\phi_p)&\tilde{C}_1\left(\frac{r_p}{f_0}\right)\cos\phi_p\\
     \tilde{C}_2\left(\frac{r_p}{f_o}\right)\sin(2\phi_p)&\tilde{C}_0\left(\frac{r_p}{f_o}\right) - \tilde{C}_2\left(\frac{r_p}{f_o}\right)\cos(2\phi_p)&\tilde{C}_1\left(\frac{r_p}{f_0}\right)\sin\phi_p\\     
     0&0&0\\     
   \end{bmatrix}
   \begin{bmatrix}
     s_x\\
     s_y\\
     s_z
   \end{bmatrix}
   \Pi\left(\frac{r_p}{2f_o\sin\alpha}\right). 
 \end{align}
 where
 \begin{alignat}{4}
   \tilde{C}_0(x) &= \frac{1}{2}(\sqrt{1 - x^2} + 1)(1 - x^2)^{-1/4} &&= 1 + \frac{x^4}{32} + \frac{x^6}{32} + \cdots,\\
   \tilde{C}_1(x) &= x(1 - x^2)^{-1/4} &&= x + \frac{x^3}{4} + \frac{5x^5}{32} + \cdots,\\
   \tilde{C}_2(x) &= \frac{1}{2}(\sqrt{1 - x^2} - 1)(1 - x^2)^{-1/4} &&= -\frac{x^2}{4} - \frac{x^4}{8} - \frac{11x^6}{128} - \cdots.
 \end{alignat}
 Similar to the monopole case, the dipole pupil function conserves power and has
 a cutoff at the objective aperture, but the dipole pupil function is
 vector-valued to model the complete electric field in the pupil plane. The
 fields in the aperture plane have a negligible $\mh{z}$ component which is a
 consequence of our assumption that the tube lens is paraxial.

 Scaling the dipole pupil function using Eq. \ref{eq:monopupildip} yields the
 dipole coherent transfer function
 \begin{align}
   \mb{C}(\taup, \so) \propto
   \begin{bmatrix}
     \tilde{C}_0\left(\frac{\text{2NA}}{n_o}\frac{\tau}{\nu_c}\right) + \tilde{C}_2\left(\frac{\text{2NA}}{n_o}\frac{\tau}{\nu_c}\right)\cos(2\phi_{\tau})&\tilde{C}_2\left(\frac{\text{2NA}}{n_o}\frac{\tau}{\nu_c}\right)\sin(2\phi_{\tau})&\tilde{C}_1\left(\frac{\text{2NA}}{n_o}\frac{\tau}{\nu_c}\right)\cos\phi_{\tau}\\
     \tilde{C}_2\left(\frac{\text{2NA}}{n_o}\frac{\tau}{\nu_c}\right)\sin(2\phi_{\tau})&\tilde{C}_0\left(\frac{\text{2NA}}{n_o}\frac{\tau}{\nu_c}\right) - \tilde{C}_2\left(\frac{\text{2NA}}{n_o}\frac{\tau}{\nu_c}\right)\cos(2\phi_{\tau})&\tilde{C}_1\left(\frac{\text{2NA}}{n_o}\frac{\tau}{\nu_c}\right)\sin\phi_{\tau}\\     
     0&0&0\\     
   \end{bmatrix}
   \begin{bmatrix}
     s_x\\
     s_y\\
     s_z
   \end{bmatrix}
   \Pi\left(\frac{\tau}{\nu_c}\right). 
 \end{align}
 We restrict our analysis to the paraxial regime by dropping all but the zero
 and first order radial terms to find that
 \begin{align}
   \mb{C}(\taup, \so) \stackrel{(p)}{\propto}
   \begin{bmatrix}
     1&0&\frac{\text{2NA}}{n_o}\frac{\tau}{\nu_c}\cos\phi_{\tau}\\
     0&1&\frac{\text{2NA}}{n_o}\frac{\tau}{\nu_c}\sin\phi_{\tau}\\
     0&0&0\\     
   \end{bmatrix}
   \begin{bmatrix}
     s_x\\
     s_y\\
     s_z
   \end{bmatrix}
   \Pi\left(\frac{\tau}{\nu_c}\right). \label{eq:coherentdip}
 \end{align}
 Under the paraxial approximation the transverse component of the dipoles
 ($s_x, s_y$) creates purely transverse fields in the pupil plane and the axial
 component of the dipole ($s_z$) creates purely radial fields in the pupil
 plane. The paraxial approximation may seem crude compared to Backer and
 Moerner's numerical results, but the approximation will allow us to calculate
 the transfer functions in closed form so that we can gain an intuition for the
 limits of the microscope. We also note that many existing works in ensemble
 polarized fluorescence microscopy make stronger approximations than ours. For
 example, Fourkas only considers the total irradiance in the back focal plane
 while ignoring the propagation of fields to the detector \cite{fourkas2001}.

 The dipole coherent spread function is the inverse Fourier transform of the
 dipole coherent transfer function. Applying Eq. \ref{eq:jincrect2} in reverse
 yields
  \begin{align}
   \mb{c}(\mb{r}, \so) \stackrel{(p)}{\propto}
   \begin{bmatrix}
     \text{jinc}_0(\nu_c r)&0&\frac{\text{NA}}{n_o}i\cos\phi\,\text{jinc}_1(\nu_c r)\\
     0&\text{jinc}_0(\nu_c r)&\frac{\text{NA}}{n_o}i\sin\phi\,\text{jinc}_1(\nu_c r)\\
     0&0&0\\     
   \end{bmatrix}
   \begin{bmatrix}
     s_x\\
     s_y\\
     s_z
   \end{bmatrix}.
  \end{align}
  Notice that the radial component of the dipole coherent spread function has a
  $\pi/2$ phase shift relative to the transverse component. This phase factor
  arises because the Fourier transform of a real and odd function is purely
  imaginary.

  \subsubsection{Paraxial dipole point spread function}
  The dipole point spread function is the (normalized) absolute square of
coherent dipole spread function
\begin{align}
  h(\mb{r}, \so) \propto \mb{c}(\mb{r}, \so)\mb{c}^\dagger(\mb{r}, \so). 
\end{align}
Plugging in the paraxial dipole coherent spread function and normalizing yields
\begin{align}
  h(\mb{r}, \so) \stackrel{(p)}{=} N\left[\text{jinc}_0^2(\nu_c r)\sin^2\vartheta + \left(\frac{\text{NA}}{n_o}\right)^2\text{jinc}_1^2(\nu_c r)\cos^2\vartheta\right],\label{eq:paradpsf}
\end{align}
where $\sin^2\vartheta = s_x^2 + s_y^2$, $\cos^2\vartheta = s_z^2$, and the normalization factor is
\begin{align}
  N = 6\nu_c^2\pi^{-3/2}\left[2 + \left(\frac{\text{NA}}{n_o}\right)^2\right]^{-1}.
\end{align}

 Notice that the transverse and radial fields are out of phase on the detector,
 so the total irradiance is the sum of the contributions from the transverse and
 radial components. In Figure \ref{fig:hdet} we plot the dipole point spread
 function for several dipole orientations dipole orientations and numerical
 apertures, and in Figure \ref{fig:microscopefig} we compare the monopole point
 spread function to the dipole point spread function. The paraxial monopole and
 dipole models are equivalent when the sample consists of only transverse
 dipoles which is clear if we notice that Eq. \ref{eq:paradpsf} reduces to an
 Airy disk when $\vartheta = \pi/2$.

\begin{figure}[h]
 \centering
   \centering
   \includegraphics[scale=0.8]{../figures/paratfs/dpsf.pdf}
   \caption{Renormalized paraxial dipole point spread function as a function of
     the scaled radial coordinate $\nu_c r$, the dipole inclination angle
     $\vartheta$, and $\text{NA}/n_o$. For small numerical apertures (left) the
     irradiance pattern created by axial dipoles (\textcolor{red}{\textbf{red}})
     is small compared to transverse dipoles (\textbf{black}), but the relative
     contribution of axial dipoles increases with the numerical aperture (see
     \textcolor{red}{\textbf{red}} lines from left to right).}
   \label{fig:hdet}
 \end{figure}

\begin{figure}[h]
 % \captionsetup{width=1.0\linewidth}
 \centering
   \centering
   \includegraphics[scale=0.8]{../figures/microscope/microscope.pdf}
   \caption{Comparison of paraxial models for monopole radiators a) and dipole
     radiators b)--d). a) Monopole radiators fill the pupil plane with a uniform
     scalar field which gives rise to an Airy disk on the detector. b) A
     transverse dipole radiator also creates an Airy disk, but the pupil plane
     is filled with a uniform vector field. c) An axial dipole radiator creates
     a radial electric field pattern in the back focal plane which creates a
     $\text{jinc}_1^2(r)$ (or higher-order Airy disk) on the detector. d)
     Dipoles that are not transverse or axial still create radially symmetric
     irradiance patterns under the paraxial approximation. Fields from
     transverse dipoles are real and even while fields from axial dipoles are
     real and odd which causes a relative $\pi/2$ phase shift for the fields on
     the detector. This phase shift means that the fields from transverse and
     axial components of the dipole do not interfere, which causes radially
     symmetric irradiance patterns.}
   \label{fig:microscopefig}
 \end{figure}

 To demonstrate the paraxial dipole point spread function we simulate a set of
 equally spaced dipoles with varying orientation
 \begin{align}
   f_{(ph1)}(r_x, r_y, \vartheta, \varphi) = \sum_{j=0}^3 \sum_{k=0}^3 \delta\left(r_x - j\right)\, \delta\left(r_y - k\right)\, \delta\left(\cos\vartheta - \cos\vartheta_j\right)\, \delta\left(\varphi - \varphi_k\right),\label{eq:phantom1}
 \end{align}
 where $\vartheta_j = j\frac{\pi}{6}$, $\varphi_k = k\frac{\pi}{4}$, the
 subscript $(ph1)$ indicates that this is the first phantom, and the spatial
 coordinates are expressed in $\mu$m. To find the irradiance pattern created by
 the phantom Eq. \ref{eq:phantom1} we plug it into Eq. \ref{eq:odpsf} and use
 the sifting property to find that
 \begin{align}
   g_{(ph1)}(r_x, r_y) = \sum_{j=0}^3 \sum_{k=0}^3 h\left(\sqrt{\left(r_x - j\right)^2 + \left(r_y - k\right)^2}, \vartheta_j\right).\label{eq:phantom1irr}
 \end{align}
 In Figure \ref{fig:ph1} we plot the phantom and scaled irradiance for an
 imaging system with $\text{NA} = 0.75$, $\lambda = 500\,\text{nm}$, and
 $n_o = 1.33$. We sample and plot the scaled irradiance at $20\times$ the
 Nyquist rate $\Delta x = 1/[20(2\nu_c)]$, so the irradiance patterns are free
 of aliasing. The output demonstrates that the irradiance pattern depends on the
 dipole inclination, but not its azimuth.

 \begin{figure}[h]
 % \captionsetup{width=1.0\linewidth}
 \centering
   \centering
   \includegraphics[scale=0.8]{../figures/paratfs/ph1.pdf}
   \caption{\textbf{Left:} A spatially and angularly sparse phantom---uniformly
     spaced single dipoles with varying orientations (increasing $\vartheta$
     from left to right and increasing $\varphi$ from bottom to top). White
     crosses mark the positions of the dipoles. \textbf{Center:} Scaled
     irradiance for an imaging system with $\text{NA} = 0.75$,
     $\lambda = 500\,\text{nm}$, and $n_o = 1.33$ sampled at $20\times$ the
     Nyquist rate. \textbf{Right:} $x$ profiles through the scaled irradiance.
     The response is independent of the azimuth angle and strongly dependent on
     the inclination angle.}
   \label{fig:ph1}
 \end{figure}

\subsection{Paraxial dipole spatial transfer function}\label{sec:trans}
The dipole spatial transfer function is the spatial Fourier transform of
the dipole point spread function (or the complex autocorrelation of the dipole
coherent transfer function). Applying the Fourier transform to Eq.
\ref{eq:paradpsf} we find that
  \begin{align}
    H(\bs{\nu}, \vartheta) \stackrel{(p)}{=} \frac{N}{\nu_c^2}\left[\text{chat}_0\left(\frac{\nu}{\nu_c}\right)\sin^2\vartheta + \left(\frac{\text{NA}}{n_o}\right)^2 \text{chat}_1\left(\frac{\nu}{\nu_c}\right)\cos^2\vartheta\right].
  \end{align}
  In Figure \ref{fig:odotf} we plot the dipole spatial transfer function for
  several dipole orientations dipole orientations and numerical apertures. We
  find that the dipole spatial transfer function is negative for axial dipoles
  at high spatial frequencies, especially for larger numerical apertures. The
  negative spatial transfer corresponds to a contrast inversion for
  high-frequency patterns of axial dipoles because the irradiance minimum
  corresponds to the position of the dipole.

\begin{figure}[h]
 % \captionsetup{width=1.0\linewidth}
 \centering
   \centering
   \includegraphics[scale=0.8]{../figures/paratfs/sdtf.pdf}
   \caption{Dipole spatial transfer function as a function of the scaled spatial
     frequency $\nu/\nu_c$, the dipole inclination angle $\vartheta$, and
     $\text{NA}/n_o$. For small numerical apertures (left) the dipole spatial
     transfer function for axial dipoles (\textcolor{red}{\textbf{red}}) is
     small compared to transverse dipoles (\textbf{black}), but the relative
     contribution of axial dipoles increases with the numerical aperture (see
     \textcolor{red}{\textbf{red}} lines from left to right). The spatial dipole
     transfer function of axial dipoles is negative at high spatial frequencies
     because the central minimum of the axial dipole point spread function
     corresponds to the position of the dipole. Equivalently, a
     high-spatial-frequency pattern of axial dipoles will generate an irradiance
     pattern where the minimum irradiance corresponds to the peak of the axial
     dipole density.}
   \label{fig:odotf}
 \end{figure}

 To demonstrate the spatial dipole transfer function we simulate a set of
 equally spaced disks with varying diameter containing fluorophores with varying
 orientation
 \begin{align}
   f_{\text{(ph2)}}(r_x, r_y, \vartheta) = \sum_{j=0}^3 \sum_{k=0}^3 \frac{1}{D_k^2}\Pi\left(\frac{1}{D_k}\sqrt{\left(r_x - j\right)^2 + \left(r_y - k\right)^2}\right) \delta\left(\cos\vartheta - \cos\vartheta_j\right)\,\label{eq:phantom2}
 \end{align}
 where $D_k = 0.15(1+k)\, \mu$m and $\vartheta_j = j\frac{\pi }{6}$. Notice that
 we have scaled the disks so that the total number of fluorophores in each disk
 is constant.

 We can calculate the irradiance pattern by taking the spatial Fourier transform
 for each orientation in the phantom, filtering the result with the dipole
 spatial transfer function, integrating over the orientations, then taking the
 inverse spatial Fourier transform. Since our phantom is angularly sparse, the
 integral becomes a sum and we find that
 \begin{align}
   g_{\text{(ph2)}}(r_x, r_y) = \mc{F}^{-1}_{\mbb{R}^2}\left\{\sum_{j}H(\nu, \vartheta_j)\mc{F}_{\mbb{R}^2}\left\{f_{\text{(ph2)}}(r_x, r_y, \vartheta_j)\right\}\right\}.
 \end{align}

 In Figure \ref{fig:ph2} we plot the phantom and scaled irradiance with the same
 imaging parameters as the previous section. The small disks create irradiance
 patterns that are similar to the point sources in the previous section, while
 larger disks create increasingly uniform irradiance patterns that hide the
 orientation of the fluorophores.

  \begin{figure}[h]
 % \captionsetup{width=1.0\linewidth}
 \centering
   \centering
   \includegraphics[scale=0.8]{../figures/paratfs/ph2.pdf}
   \caption{\textbf{Left:} A spatially dense and angularly sparse
     phantom---uniformly spaced disks with varying size (increasing $D$ from
     bottom to top) and dipole orientation (increasing $\vartheta$ from left to
     right) \textbf{Center:} Scaled irradiance for an imaging system with
     $\text{NA} = 0.75$, $\lambda = 500\,\text{nm}$, and $n_o = 1.33$ sampled at
     $20\times$ the Nyquist rate. \textbf{Right:} $x$ profiles through the
     scaled irradiance. Larger disks generate increasingly uniform irradiance
     patterns with fewer details that may indicate the orientation of
     fluorophores.}
   \label{fig:ph2}
 \end{figure}

 \subsubsection{Paraxial dipole angular transfer function}
 To calculate the angular dipole transfer function we take the spherical Fourier
 transform of the dipole point spread function
 \begin{align}
H_l^m(\mb{r}) \stackrel{}{=} \int_{\mbb{S}^2}d\so\, h(\mb{r}, \so)Y_l^{m*}(\so).
 \end{align}
 After evaluating the integrals and normalizing, the angular dipole transfer
 function is
 \begin{align}
   H_l^m(\mb{r}) \stackrel{(p)}{=} &\frac{N}{3}\left[2\text{jinc}_0^2(\nu_cr) + \left(\frac{\text{NA}}{n_o}\right)^2\text{jinc}_1^2(\nu_c r)\right]\Lambda_0\delta_{\ell0}\delta_{m0} + \nonumber\\& \frac{N}{3}\left[-2\text{jinc}_0^2(\nu_c r) + 2\left(\frac{\text{NA}}{n_o}\right)^2\text{jinc}_1^2(\nu_c r)\right]\Lambda_2\delta_{\ell2}\delta_{m0},
 \end{align}
where $\Lambda_{\ell} = \sqrt{4\pi/(2\ell + 1)}$.

In Figure \ref{fig:atf} we plot the dipole angular transfer function for both
spherical harmonic terms and several numerical apertures. Note that the dipole
angular transfer function can be negative because the spherical harmonics
contain negative values. The $\ell=0$ term shows that angularly uniform
distributions of dipole create spatial irradiance patterns that are nearly
identical to the Airy disk, while the $\ell=2$ term shows a negative pattern
because of the large contribution of the transverse negative dipoles in the
$Y_2^0$ spherical harmonic.

 % The degree $\ell$ and order $m$ of the non-zero terms in the angular dipole
 % transfer function . First, the spatio-angular point spread function is symmetric
 % under inversion of the dipole moment. This symmetry is maintained in the
 % angular transfer function since all of the non-zero terms have an even degree
 % $\ell=0, 2$, and the corresponding spherical harmonics $Y_0^0(\so)$,
 % $Y_2^0(\so)$ are symmetric under inversion. Second, we showed that under the
 % paraxial approximation our kernel is rotationally symmetric which allows us to
 % write the kernel as a function of $r$ instead of $\mb{r}$. This symmetry is
 % maintained in the angular dipole transfer function since all of the non-zero
 % terms have order $m=0$. Finally, the dipole point spread function consists of
 % degree-2 trigonometric polynomials which implies that the dipole angular
 % transfer function will only contain $\ell=0, 2$ terms.

\begin{figure}[h]
 \centering
   \centering
   \includegraphics[scale=0.8]{../figures/paratfs/adtf.pdf}
   \caption{Paraxial dipole angular transfer function in terms of a scaled
     radial detection coordinate $\nu_c r$, the spherical harmonic degree
     $\ell$, and $\text{NA}/n_o$. Angularly uniform distributions of dipoles
     $\ell=0$ generate a spatial pattern that is nearly identical to an Airy
     disk. $\ell=2$ distributions have a negative response because
     $Y_2^0(\mh{s})$ is negative for transverse directions. As the numerical
     aperture increases, the relative contribution of positive axial dipoles in
     the $\ell=2$ distribution increases.}
   \label{fig:atf}
 \end{figure}

 To demonstrate the dipole angular transfer function we simulate a set of
 equally spaced fluorophore distributions with varying orientation and angular
 distributions
 \begin{align}
   f_{\text{(ph3)}}(r_x, r_y, \vartheta) = \sum_{j=0}^3 \sum_{k=0}^3 \delta\left(r_x - j\right)\, \delta\left(r_y - k\right)\, f_{\text{(cone)}}\left(\vartheta, \varphi; \vartheta_j, 0, \Delta_k\right),\label{eq:phantom3}
 \end{align}
 where
 \begin{align}
   f_{\text{(cone)}}(\so; \so', \Delta) = f_{\text{(cone)}}(\vartheta, \varphi; \vartheta', \varphi', \Delta) = \frac{1}{2\pi(1 - \cos\Delta)}\Pi\left(\frac{\mh{s}\cdot\mh{s}'}{\cos\Delta}\right),
 \end{align}
 is an angular double cone distribution with central direction $\mh{s}'$ and
 cone half-angle $\Delta$; $\vartheta_j = j\frac{\pi}{6}$; and
 $\Delta_k = k\frac{\pi}{6}$. Notice that when $\Delta = 0$ the angular double
 cone reduces to a single direction, and when $\Delta = \pi/2$ the angular
 double cone reduces to an angularly uniform distribution.

 Our first step towards the irradiance pattern is to calculate the dipole
 angular spectrum of the phantom. In Appendix \ref{sec:cone} we calculate the
 spherical Fourier transform of the double cone distribution
 $F^m_{\ell,\text{(cone)}}(\vartheta', \varphi'; \Delta)$ which we can use to
 express the dipole angular spectrum as
 \begin{align}
   F^{m}_{\ell,\text{(ph3)}}(r_x, r_y, \vartheta) = \sum_{j=0}^3 \sum_{k=0}^3 \delta\left(r_x - j\right)\, \delta\left(r_y - k\right)\, F^m_{\ell,\text{(cone)}}\left(\vartheta_j, 0, \Delta_k\right).\label{eq:phantom3spect}
 \end{align}

 To find the irradiance we plug the dipole angular spectrum into Eq.
 \ref{eq:atf-form} and use the sifting property we find that
 \begin{align}
   g_{\text{(ph3)}}(r_x, r_y) = \sum_{\ell m}\sum_{j=0}^3 \sum_{k=0}^3 H_\ell^m\left(\sqrt{\left(r_x - j\right)^2 + \left(r_y - k\right)^2}\right) F^m_{\ell,\text{(cone)}}\left(\vartheta_j, 0, \Delta_k\right).\label{eq:phantom3irr}
 \end{align}

 In Figure \ref{fig:ph3} we plot the phantom and scaled irradiance with the same
 imaging parameters as the previous sections. For small cone angles the
 irradiance patterns are similar to the point sources in the previous sections,
 while larger cone angles create increasingly uniform irradiance patterns that
 hide the angular information about the distributions.

  \begin{figure}[h]
 % \captionsetup{width=1.0\linewidth}
 \centering
   \centering
   \includegraphics[scale=0.8]{../figures/paratfs/ph3.pdf}
   \caption{\textbf{Left:} A spatially sparse and angularly dense
     phantom---uniformly spaced double cone distributions of fluorophores with
     varying central direction (increasing $\vartheta'$ from left to right) and
     varying cone half-angle (increasing $\Delta$ from bottom to top).
     \textbf{Center:} Scaled irradiance for an imaging system with
     $\text{NA} = 0.75$, $\lambda = 500\,\text{nm}$, and $n_o = 1.33$ sampled at
     $20\times$ the Nyquist rate. \textbf{Right:} $x$ profiles through the
     scaled irradiance. Small cone angles have irradiance patterns that vary
     with the central direction, while larger cones angles have increasingly
     uniform irradiance patterns that hide angular information.}
   \label{fig:ph3}
 \end{figure}

 \subsubsection{Paraxial dipole spatio-angular transfer function}
 We can calculate the dipole spatio-angular transfer function by taking the
 spatial Fourier transform of the dipole angular transfer function (or the
 spherical Fourier transform of the dipole spatial transfer function) to find
 that
 \begin{align}
H_{\ell}^m(\mb{r}) \stackrel{(p)}{=} &\frac{N}{3\nu_c^2}\left[2\text{chat}_0\left(\frac{\nu}{\nu_c}\right) + \left(\frac{\text{NA}}{n_o}\right)^2\text{chat}_1\left(\frac{\nu}{\nu_c}\right)\right]\Lambda_0\delta_{\ell0}\delta_{m0}+\nonumber\\ &\frac{N}{3\nu_c^2}\left[-2\text{chat}_0\left(\frac{\nu}{\nu_c}\right) + 2\left(\frac{\text{NA}}{n_o}\right)^2\text{chat}_1\left(\frac{\nu}{\nu_c}\right)\right]\Lambda_2\delta_{\ell2}\delta_{m0}.
 \end{align}

 In Figure \ref{fig:dsatf} we plot the dipole spatio-angular transfer function
 for both spherical harmonic terms and several numerical apertures. The $\ell=0$
 term shows that angularly uniform distributions of dipole create spatial
 irradiance patterns that are nearly identical to the Airy disk, with high
 frequencies increasingly suppressed as the numerical aperture increases. The
 $\ell=2$ term shows a negative pattern because of the large contribution of the
 transverse negative dipoles in the $Y_2^0$ spherical harmonic, and as the
 numerical aperture increases the relative contribution of the positive axial
 dipoles increases and the $\ell=2$ term becomes less negative.

\begin{figure}[h]
 \centering
   \centering
   \includegraphics[scale=0.8]{../figures/paratfs/satf.pdf}
   \caption{Spatio-angular dipole transfer function as a function of the scaled
     spatial frequency $\nu/\nu_c$, the spherical harmonic degree $\ell$, and
     $\text{NA}/n_o$. When the numerical aperture is small the transverse
     dipoles contribute the most to the signal which gives rise to a positive
     $\ell=0$ component and a negative $\ell=2$ component. As the numerical
     aperture increases, the relative contribution of axial dipoles increases
     and the $\ell=2$ component becomes less negative.}
   \label{fig:dsatf}
 \end{figure}

 To demonstrate the spatio-angular transfer function, we simulate a set of
 equally spaced disks of fluorophores with varying radius and angular
 distributions
 \begin{align}
   f_{\text{(ph4)}}(r_x, r_y, \vartheta, \varphi) = \sum_{j=0}^3 \sum_{k=0}^3 \frac{1}{D_k^2}\Pi\left(\frac{1}{D_k}\sqrt{\left(r_x - j\right)^2 + \left(r_y - k\right)^2}\right)f_{\text{(cone)}}\left(\vartheta, \varphi; \frac{\pi}{2}, 0, \Delta_j\right),\label{eq:phantom4}
 \end{align}
 where $D_k = 0.15(1+k)\, \mu$m, and $\Delta_j = j\frac{\pi}{6}$.

 Our first step towards calculating the irradiance pattern is to calculate the
 dipole spatio-angular spectrum given by the spatial Fourier transform of the dipole angular spectrum  
  \begin{align}
   \mathsf{F}_{\ell,\text{(ph4)}}^m(\nu_x, \nu_y) = \mc{F}_{\mbb{R}^2}\left\{\sum_{j=0}^3 \sum_{k=0}^3 \frac{1}{D_k^2}\Pi\left(\frac{1}{D_k}\sqrt{\left(r_x - j\right)^2 + \left(r_y - k\right)^2}\right)F_{\ell,\text{cone}}^m\left(\frac{\pi}{2}, 0, \Delta_j\right)\right\}.
  \end{align}
  We can plug the dipole spatio-angular spectrum into Eq. \ref{eq:saft} to find
  the scaled irradiance as 
  \begin{align}
   g_{\text{(ph4)}}(r_x, r_y) = \mc{F}^{-1}_{\mbb{R}^2}\left\{\sum_{\ell m}\mathsf{H}_{\ell}^m(\nu_x, \nu_y)\mathsf{F}_{\ell,\text{(ph4)}}^m(\nu_x, \nu_y)\right\}.
  \end{align}

  In Figure \ref{fig:ph4} we plot the phantom and scaled irradiance with the
  same imaging parameters as the previous sections. Small cone angles and small
  disks create relatively unique irradiance patterns, while increasing the cone
  angle or disk size creates increasingly irradiance patterns.

  \begin{figure}[h]
 % \captionsetup{width=1.0\linewidth}
 \centering
   \centering
   \includegraphics[scale=0.8]{../figures/paratfs/ph4.pdf}
   \caption{\textbf{Left:} A spatially and angularly dense phantom---uniformly
     spaced disks with varying size (increasing $D$ from bottom to top) and
     double cone half angle (increasing $\Delta$ from left to right)
     \textbf{Center:} Scaled irradiance for an imaging system with
     $\text{NA} = 0.75$, $\lambda = 500\,\text{nm}$, and $n_o = 1.33$ sampled at
     $20\times$ the Nyquist rate. \textbf{Right:} $x$ profiles through the
     scaled irradiance.}
   \label{fig:ph4}
 \end{figure}

\section{Discussion}\label{sec:discussion}
\subsection{Reconstructions}
% Since the dipole spatio-angular transfer function effectively contains it is
% tempting to conclude that fluorescence microscopes pass three times as much
% information about the dipoles as monopoles. This is not true. The microscope
% still only measures a single member of $\mbb{L}_2(\mbb{R}^2)$, so at best we can
% recover the same amount of information about the object. In other words, we know
% immediately that the microscope is a rank-1 operator since it only measures one
% continuous function.
\begin{itemize}
\item Very ill-posed reconstruction problem by dimensionality.
\item Better-posed reconstruction problem by adding polarized illumination/detection
  formulated as $\mc{H}: \mbb{L}_2(\mbb{R}^2\times\mbb{S}^2) \rightarrow \mbb{L}_2(\mbb{R}^3\times\mbb{S}^1)$ and/or multiple views $\mc{H}: \mbb{L}_2(\mbb{R}^2\times\mbb{S}^2) \rightarrow \mbb{L}_2(\mbb{R}^2\times\mbb{S}^1)^V$.
\item Mention SVD for future papers
\end{itemize}

% TODO Make clear that a later paper will include a full discussion of SVD
% to find the object space singular functions that span the measurement space of
% the instrument.

\subsection{Alternative transfer functions}
Throughout this work we have used the spherical harmonic functions as a basis
for functions on the sphere, but there are other basis functions that can be
advantageous in some cases. Backer and [TODO Brasselet, others] have used
the second moments as basis functions for the sphere because they arise
naturally when computing the dipole point spread function. Mathematically,
Backer and Moerner \cite{backer2014} use an alternative to the dipole angular
transfer function that uses the second moments as basis functions so their
forward model becomes
\begin{align}
  g(\rd) = \sum_{j=1}^6 H_j(\rd - \ro)F_j(\ro),
\end{align}
where
\begin{align}
  H_j(\rd - \ro) &= \int_{\mbb{S}^2}d\so\, h(\rd - \ro, \so)Z_j(\so),\\
  F_j(\ro) &= \int_{\mbb{S}^2}d\so\, f(\ro, \so)Z_j(\so),
\end{align}
and $Z_j(\mh{s}) = \{s_x^2, s_y^2, s_z^2, s_xs_y, s_ys_z, s_xs_z\}$ are the
second moments. This formulation is similar to the angular transfer function
approach because it exploits the spatial sparsity of sample, but it does not
require a cumbersome expansion of the dipole point spread function onto
spherical harmonics.

However, the spherical harmonics provide several advantages over the second
moments. First, the spherical harmonics form a complete basis for functions on
the sphere, while the second moments span a much smaller function space. The
usual approach to extending the span of the second moments is to use the fourth
(or higher) moments, but these are completely new basis functions. Second, the
spherical harmonics are orthonormal which will allow us to easily deploy
invaluable tools from linear algebra---linear subspaces, rank, decompositions,
etc.--- to analyze and compare microscope designs. Finally, using the spherical
harmonics provides access to a set of fast algorithms. The naive expansion of an
arbitrary discretized $N$ point spherical function onto spherical harmonics (or
second moments) requires a $\mc{O}(N^2)$ matrix multiplication, while pioneering
work by Driscoll and Healy [cite] showed that the forward discrete fast
spherical harmonic transform can be computed with a $\mc{O}(N(\log N)^2)$
algorithm and its inverse can be computed with a $\mc{O}(N^{3/2})$ algorithm. To
our knowledge no similarly fast algorithms exist for expansion onto the
higher-order moments.

TODO: Cite Tournier/Descoteux for an easy change of basis between spherical
harmonics and second moments

\subsection{What determines the angular bandwidth?}
Spatial imaging systems have a spatial bandwidth which characterizes is the
highest spatial frequency that the system can image. Similarly, angular imaging
systems have an angular bandwidth that characterizes the highest angular
frequency, but in the angular case there are two different types of angular
bandwidths that we call $ell$ and $m$ bandwidths. The $\ell$-bandwidth can be
interpreted in a similar way to the spatial bandwidth---it characterizes the
smallest angular features that the imaging system can measure. The $m$-bandwidth
does not have a direct analog in the spatial domain---it characterizes the
angular uniformity of the imaging system. If the $\ell$ and $m$ bandwidths are
equal then the imaging system can be said to be $\textit{rotation invariant}$.

The spatial bandwidth of a fluorescence microscope is well known to be
$\nu_c = \frac{2\text{NA}}{\lambda}$. In other words, we can increase the
spatial resolution of a fluorescence microscope by increasing the NA of the
instrument or by choosing a fluorophore with a shorter emission wavelength.
Similarly, the angular bandwidth of a fluorescence microscope depends on both
the instrument and the sample.

The microscope we considered in this work has an $ell$ bandwidth of $\ell = 2$
and an $m$ bandwidth of $m_c = 0$, so it is not a rotation invariant imaging
system. In future work we will consider several approaches to improving the
angular bandwidths in detail, but we briefly mention that non-paraxial
microscopes, microscopes with polarizers on the illumination or detection paths,
and multiview microscopes all have higher angular bandwidths that the microscope
considered here.

The angular bandwidth is also sample dependent. Even though they are just a
model, monopoles emit light isotropically so they have an $\ell$ bandwidth of 1,
while dipoles have an $\ell$ bandwidth of 2, and higher-order excitation and
detection moments will have even higher bandwidths. Multi-photon excitation and
other non-linear methods can also increase the $\ell$ bandwidth.

\subsection{Towards more realistic models}
The theoretical model we presented in this work is an extreme simplification of a
real microscope. We have ignored the effects of thick samples, refractive-index
mismatch, aberration, scattering, rotation of fluorophores, finite fluorescence
lifetimes, and interactions between fluorophores among others. Because of this
long list of unknown effects, real experiments will likely require extensions of
the models developed here.

TODO: Discuss/plan future papers. SVD? Defocus/3D imaging:
$\mc{H}: \mbb{L}_2(\mbb{R}^3\times\mbb{S}^2) \rightarrow \mbb{L}_2(\mbb{R}^3)$.

TODO: Mention and cite rotating dipoles like Lew (model start point, rotation, end point):
$\mc{H}: \mbb{L}_2(\mbb{R}^3\times\mbb{S}^2\times\mbb{SO}^3\times\mbb{S}^2) \rightarrow \mbb{V}$.

TODO: Cite connection between dipole pupil and Jones pupil. We can use the
Jones/dipole pupil as a way to extend our model in a straightforward way. Add
phase aberration (like the Gibson-Lanni model), polarization aberrations,
amplitude masks, phase masks, etc.

TODO: Cite Piestun Green's tensor engineering.

\section{Conclusion}
Most models of fluorescence microscopes use a monopole model to describe the
object and the scalar approximation to model the imaging system. A more complete
description of any fluorescence microscope requires a dipole model of the object
and vector optics to model the propagation of light through the microscope. We
developed several transfer functions that simplify the mapping between the
dipole density and the irradiance pattern on the detector, and we demonstrated
these transfer functions by efficiently simulating a paraxial fluorescence
microscope.

\bibliography{paper}
\appendix
\section{Spherical harmonics}\label{sec:sph}
The spherical harmonic function of degree $\ell$ and order $-\ell \leq m \leq \ell$
is defined as \cite{schaeffer2013}
\begin{align}
Y_\ell^m(\vartheta, \varphi) = \sqrt{\frac{2\ell+1}{4\pi}}\sqrt{\frac{(\ell-|m|)!}{(\ell+|m|)!}}P_\ell^m\left(\cos\vartheta\right)\exp(i m \varphi),
\end{align}
where $P_\ell^m(\cos\theta)$ are the associated Legendre polynomials with the
Condon-Shortley phase
\begin{align}
  P_\ell^m(x) = (-1)^m(1-x^2)^{|m|/2}\frac{d^{|m|}}{dx^{|m|}}P_\ell(x),
\end{align}
and $P_\ell(x)$ are the Legendre polynomials defined by the recurrence
\begin{align}
  P_0(x) &= 1,\\
  P_1(x) &= x,\\
  \ell P_\ell(x) &= (2\ell-1)xP_{\ell-1}(x) - (\ell-1)P_{\ell-2}(x). 
\end{align}
The spherical harmonics are orthonormal, which means that
\begin{align}
  \int_{\mbb{S}^2}d\mh{s}\, Y_\ell^m(\mh{s}){Y}_{\ell'}^{m'*}(\mh{s}) = \delta_{\ell\ell'}\delta_{mm'},
\end{align}
where $\delta_{\ell\ell'}$ denotes the Kronecker delta. The spherical harmonics form a
complete basis so an arbitrary function on the sphere $f(\mh{s})$ can be
expanded into a sum of weighted spherical harmonic functions
\begin{align}
  f(\mh{s}) = \sum_{\ell=0}^{\infty}\sum_{m=-\ell}^{l}F_\ell^mY_\ell^m(\mh{s}).
\end{align}
We can find the spherical harmonic coefficients $F_\ell^m$ for a given function
using Fourier's trick---multiply both sides by $Y_\ell^{m*}(\mh{s})$,
integrate over the sphere, and exploit orthogonality to find that
\begin{align}
  F_\ell^m = \int_{\mbb{S}^2}d\mh{s}\, f(\mh{s})Y_\ell^{m*}(\mh{s}).
\end{align}
The coefficients $F_\ell^m$ are called the \textit{spherical Fourier transform}
of a spherical function.

We briefly show how two properties of spherical functions propagate to the
spherical Fourier transform. First, the spherical harmonic coefficients of
spherical functions that can be written entirely as a function of the
inclination angle $\vartheta$ are zero when $m\neq0$ because
\begin{align}
  \int_{\mbb{S}^2}d\mh{s}\, f(\vartheta)Y_\ell^{m*}(\mh{s}) &= \int_0^{2\pi}d\phi\, \text{exp}[-im\varphi]\int_0^{\pi}d\vartheta\, \sin\vartheta f(\vartheta)P_\ell^m(\cos\vartheta),\\ &= \delta_{m0}\int_0^{\pi}d\vartheta\, \sin\vartheta f(\vartheta)P_\ell^m(\cos\vartheta).
\end{align}
Second, the spherical harmonic coefficients of spherical functions that are
symmetric under inversion $f(\mh{s}) = f(-\mh{s})$ are zero when $\ell$ is odd
because $Y_\ell^m(-\so) = (-1)^\ell Y_\ell^m(\mh{s})$, so if $\ell$ is odd then
\begin{align}
  \int_{\mbb{S}^2}d\mh{s}\, f(\mh{s})Y_{2n+1}^{m*}(\mh{s}) &= \int_{\mbb{S}^2/2}d\mh{s}\, f(+\mh{s})Y_{2n+1}^{m*}(+\mh{s}) + \int_{\mbb{S}^2/2}d\mh{s}\, f(-\mh{s})Y_{2n+1}^{m*}(-\mh{s}),\\
&=\int_{\mbb{S}^2/2}d\mh{s}\, f(\mh{s})Y_{2n+1}^{m*}(+\mh{s}) - \int_{\mbb{S}^2/2}d\mh{s}\, f(\mh{s})Y_{2n+1}^{m*}(\mh{s}) = 0.                                                                
\end{align}
\section{Relationships between special functions} \label{sec:special}
Our first task is to show that 
\begin{align}
  i^{n}\left\{\substack{
  \text{exp}(in\phi_r)\\
    \cos(n\phi_r)\\
    \sin(n\phi_r)
  }\right\}\text{jinc}_{n}(r) &\stackrel{\mathcal{F}_{\mbb{R}^2}}{\longrightarrow}   (2\nu)^{n}\left\{\substack{
    \text{exp}(in\phi_{\nu})\\
    \cos(n\phi_{\nu})\\
    \sin(n\phi_{\nu})
  }\right\}\Pi(\nu). \label{eq:jincrect3}
  \end{align}
  Writing the inverse Fourier transform in polar coordinates yields
  \begin{align}
    =2^n\int_0^{1/2}d\nu\, \nu^{n+1}\int_0^{2\pi}d\phi_\nu\left\{\substack{
    \text{exp}(in\phi_{\nu})\\
    \cos(n\phi_{\nu})\\
    \sin(n\phi_{\nu})
  }\right\}\text{exp}[2\pi i\nu r\cos(\phi_\nu - \phi_r)].
  \end{align}
  The azimuthal integral can be evaluated in terms of an $n^{th}$ order Bessel
  function (see Barrett and Myers 4.111 for the complex case \cite{barrett2004})
  \begin{align}
    =2^n 2\pi i^n \left\{\substack{
    \text{exp}(in\phi_{r})\\
    \cos(n\phi_r)\\
    \sin(n\phi_r)
  }\right\}\int_0^{1/2}d\nu\, \nu^{n+1}J_n(2\pi\nu r).
  \end{align}
  We can use the following identity (see GR 6.561-5 \cite{gradshteyn2007})
  \begin{align}
    \int_0^1 du\, u^{n + 1}J_{n}(au) = a^{-1}J_{n + 1}(a)
  \end{align}
  with a change of variable $u = 2\nu$ to find the final result
  \begin{align}
    &=2^n 2\pi i^n \left\{\substack{
    \text{exp}(in\phi_{r})\\
    \cos(n\phi_r)\\
    \sin(n\phi_r)
  }\right\}\int_0^{1}\frac{du}{2}\, \left(\frac{u}{2}\right)^{n+1}J_n(\pi u r)
    =i^n \left\{\substack{
    \text{exp}(in\phi_{r})\\
    \cos(n\phi_r)\\
    \sin(n\phi_r)
  }\right\}\frac{J_{n+1}(\pi r)}{2r}
    =i^n \left\{\substack{
    \text{exp}(in\phi_{r})\\
    \cos(n\phi_r)\\
    \sin(n\phi_r)
  }\right\}\text{jinc}_{n}(r).
  \end{align}

\begin{figure}[h]
 \centering
   \centering
   \includegraphics[width = 0.45\textwidth]{../figures/special/special.pdf}
   \caption{The relationships between special functions. The chat functions are
     defined as the two-dimensional Fourier transform of the squared jinc
     functions, and they can be calculated with the two-dimensional complex
     autocorrelations (denoted by $\star_2$) of the complex-weighted rectangle
     functions.}
   \label{fig:special}
 \end{figure}

We can use the relationship in Eq. \ref{eq:jincrect3} to express the chat
functions in terms of a complex autocorrelation---see the diagram in Figure \ref{fig:special}. Starting with the definition of
the $n^{th}$-order chat function
\begin{align}
  \text{chat}_{n}(\nu) = \int_{\mbb{R}^2}d\mb{r}\, \text{jinc}_{n}^2(|\mb{r}|)\, \text{exp}\left[-2\pi i\mb{r}\nu\right],  \label{eq:ftA}
\end{align}
we can rewrite the integrand in terms of the absolute square of a simpler
function with a known Fourier transform
\begin{align}
  \text{chat}_{n}(\nu) = \int_{\mbb{R}^2}d\mb{r}\, |t_n(\mb{r})|^2\, \text{exp}\left[-2\pi i\mb{r}\nu\right].  \label{eq:ftB}
\end{align}
\begin{align}
  t_n(\mb{r}) = i^n\text{exp}[in\phi_r]\text{jinc}_{n}(r).
\end{align}
Now we can apply the autocorrelation theorem to rewrite the Fourier transform as
\begin{align}
  \text{chat}_{n}(\nu) = \int_{\mbb{R}^2}d\bs{\tau}\, T_n(\bs{\tau})T^*_n(\bs{\tau} - \nu), \label{eq:autocorr}
\end{align}
where the function to be autocorrelated can be found with the help of Eq. \ref{eq:jincrect3}
\begin{align}
  T_n(\bs{\tau}) = \int_{\mbb{R}^2}d\mb{r}\, t_n(\mb{r})\text{exp}\left[-2\pi i\mb{r}\cdot\bs{\tau}\right] = (2\tau)^n\text{exp}[in\phi_\tau]\Pi\left(\tau\right). \label{eq:tau}
\end{align}
It will be more convenient to set up the autocorrelation in Cartesian coordinates 
\begin{align}
  T_n(\bs{\tau}) &= 2^n(\tau_x + i\tau_y)^{n}\Pi\left(\sqrt{\tau_x^2 + \tau_y^2}\right). \label{eq:tn}
\end{align}
Plugging Eq. \ref{eq:tn} into Eq. \ref{eq:autocorr} gives
\begin{align}
  \text{chat}_{n}(\nu) = 4^n\int_{\mbb{R}^2}d\bs{\tau}\, (\tau_x^2 + \tau_y^2 - \nu\tau_x)^{n}\Pi\left(\sqrt{\tau_x^2 + \tau_y^2}\right)\Pi\left(\sqrt{(\tau_x - \nu)^2 + \tau_y^2}\right).
\end{align}
We can interpret the autocorrelation as an integral over a region of overlap
between a circle centered at the origin and a circle shifted to the right by
$\nu$ (a geometric lens). Using the construction in Figure \ref{fig:geometry} we
can express this region as
\begin{align}
  \text{chat}_{n}(\nu) = 4^{n+1}\Bigg[&\int_0^{1/2}\tau d\tau\int_0^{\cos^{-1}\nu}d\phi_{\tau}(\tau^2 - \nu\tau\cos\phi_{\tau})^{n} -\nonumber \\ &\int_{0}^{\nu/2}d\tau_x\int_0^{\frac{\tau_x}{\nu}\sqrt{1 - \nu^2}}d\tau_y(\tau_x^2 + \tau_y^2 - \nu\tau_x)^{n}\Bigg]\Pi\left(\frac{\nu}{2}\right).
\end{align}
\begin{figure}[h]
 \centering
   \centering
   \includegraphics[width = 1.0\textwidth]{../figures/autocorrelation/autocorrelation.pdf}
   \caption{Geometric construction for evaluating the autocorrelation. We need
     to integrate over the overlapping region of two circles with radius $1/2$
     and distance $\nu$ between their centers. The region is four times the
     difference in area between a sector of angle $\arccos(\nu)$ and radius
     $1/2$ and a right triangle with base $\nu/2$ and hypotenuse $1/2$.}
   \label{fig:geometry}
 \end{figure}

For $n=0$:
\begin{align}
  \text{chat}_0(\nu) &= 4\left[\int_0^{1/2}\tau d\tau\int_0^{\cos^{-1}\nu}d\phi_{\tau} - \int_{0}^{\nu/2}d\tau_x\int_0^{\frac{\tau_x}{\nu}\sqrt{1 - \nu^2}}d\tau_y\right]\Pi\left(\frac{\nu}{2}\right),\\
  \text{chat}_0(\nu) &= \frac{1}{2}\left[\cos^{-1}\nu - \nu\sqrt{1 - \nu^2}\right]\Pi\left(\frac{\nu}{2}\right),
\intertext{which is a well-known result \cite{goodman1996, mertz2009, bracewell2004}. For $n=1$:}
  \text{chat}_1(\nu) &= 16\Bigg[\int_0^{1/2}\tau d\tau\int_0^{\cos^{-1}\nu}d\phi_{\tau}(\tau^2 - \nu\tau\cos\phi_{\tau}) - \\ &\qquad\quad\int_{0}^{\nu/2}d\tau_x\int_0^{\frac{\tau_x}{\nu}\sqrt{1 - \nu^2}}d\tau_y\, (\tau_x^2 + \tau_y^2 - \nu\tau_x)\Bigg]\Pi\left(\frac{\nu}{2}\right),\\
  \text{chat}_1(\nu) &= \frac{1}{2}\left[\cos^{-1}\nu - \nu (3 - 2\nu^2)\sqrt{1 - \nu^2}\right]\Pi\left(\frac{\nu}{2}\right). 
\end{align}

\section{Spherical Fourier transform of a double cone}\label{sec:cone}
In this appendix we will evaluate the spherical Fourier transform of a
normalized double-cone angular distribution with central direction $\mh{s}'$ and
cone half-angle $\Delta$
\begin{align}
  f_{\text{(cone)}}(\mh{s}; \mh{s}', \Delta) = \frac{1}{2\pi(1 - \cos\Delta)}\Pi\left(\frac{\mh{s}\cdot\mh{s}'}{\cos\Delta}\right). 
\end{align}
The spherical Fourier transform is
\begin{align}
  F_{\ell\text{(cone)}}^m(\mh{s}', \Delta) = \int_{\mbb{S}^2}d\mh{s}\, f_{\text{(cone)}}(\mh{s}; \mh{s}', \Delta)Y_\ell^{m*}(\mh{s}).
\end{align}
The limits of integration will be difficult to find unless we change coordinates
to exploit the axis of symmetry $\mh{s}'$. Since the spherical function is
rotationally symmetric about $\mh{s}'$ we can rotate the function so that the axis of
symmetry is aligned with $\mh{z}$ and multiply by $\sqrt{\frac{4\pi}{2l+1}}Y_\ell^m(\mh{s}')$ to account for the rotation
\begin{align}
    F_{\ell\text{(cone)}}^m(\mh{s}', \Delta) = \sqrt{\frac{4\pi}{2l+1}}Y_\ell^m(\mh{s}')\int_{\mbb{S}^2}d\mh{s}\, f_{\text{(cone)}}(\vartheta; \mh{z}, \Delta)Y_\ell^0(\mh{s}). 
\end{align}
In this coordinate system the double cone is independent of the azimuthal angle,
so we can evaluate the azimuthal integral and express the function in terms of an
integral over
$\vartheta$
\begin{align}
    F_{\ell\text{(cone)}}^m(\mh{s}', \Delta) = 2\pi Y_\ell^m(\mh{s}')\int_{0}^\pi d\vartheta\, \sin\vartheta f_{\text{(cone)}}(\vartheta; \mh{z}, \Delta)P_\ell(\cos\vartheta). 
\end{align}
The function $f_{\text{(cone)}}(\vartheta; \mh{z}, \Delta)$ is only non-zero on
the intervals $\vartheta \in [0, \Delta]$ and
$\vartheta \in [\pi - \Delta, \pi]$ so
\begin{align}
    F_{\ell\text{(cone)}}^m(\mh{s}', \Delta) = \frac{Y_\ell^m(\mh{s}')}{2(1 - \cos\Delta)}\left[\int_{0}^\Delta d\vartheta\, \sin\vartheta P_\ell(\cos\vartheta) + \int_{\pi - \Delta}^\pi d\vartheta\, \sin\vartheta P_\ell(\cos\vartheta)\right]. 
\end{align}
Applying a change of coordinates with $u = \cos\vartheta$ yields
\begin{align}
    F_{\ell\text{(cone)}}^m(\mh{s}', \Delta) = \frac{Y_\ell^m(\mh{s}')}{2(1 - \cos\Delta)}\left[\int_{\cos\Delta}^1 d\vartheta\, P_\ell(u) + \int_{-1}^{-\cos\Delta} d\vartheta\, P_\ell(u)\right]. 
\end{align}
The Legendre polynomials $P_\ell(u)$ are even (odd) on the interval [-1, 1] when
$\ell$ is even (odd), so the pair of integrals will be identical when $\ell$ is
even and cancel when $\ell$ is odd. For even $\ell$
\begin{align}
  F_{\ell\text{(cone)}}^m(\mh{s}', \Delta) = \frac{Y_\ell^m(\mh{s}')}{1 - \cos\Delta}\int_{\cos\Delta}^1 d\vartheta\, P_\ell(u).
\end{align}
The integral evaluates to (see GR 7.111 \cite{gradshteyn2007})
\begin{align}
  \int_{\cos\Delta}^1 d\vartheta\, P_\ell(u) =
\begin{cases}
  1 - \cos\Delta\,, &\ell = 0\,,\\
  \sin\Delta\, P_l^{-1}(\cos\Delta)\,, &\text{else},
\end{cases}
\end{align}
where $P_l^{-1}(\cos\Delta)$ is the associated Legendre polynomial with order
$m=-1$, not an inverse Legendre polynomial. Bringing everything together

\begin{align}
  F_{\ell\text{(cone)}}^m(\mh{s}', \Delta) =
  \begin{cases}
    \sqrt{1/(4\pi)}, & \ell = 0,\\
    0, & \ell\, \text{odd},\\
    Y_\ell^m(\mh{s}')\cot\left(\frac{\Delta}{2}\right)P_l^{-1}(\cos\Delta), & \ell > 0\,\, \text{even}.
  \end{cases}
\end{align}

\end{document}