### 2019-02-21 notes/2019-01-28-angular-optimize/report/report.tex
Latest angular notes with summary.

\documentclass[11pt]{article}

%%%%%%%%%%%%
% Packages %
%%%%%%%%%%%%
\usepackage[dvipsnames]{xcolor}
\hyphenpenalty=10000
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}

\usepackage{tocloft}
\usepackage[linesnumbered,ruled]{algorithm2e}
\renewcommand\cftsecleader{\cftdotfill{\cftdotsep}}
\def\undertilde#1{\mathord{\vtop{\ialign{##\crcr
$\hfil\displaystyle{#1}\hfil$\crcr\noalign{\kern1.5pt\nointerlineskip}
$\hfil\tilde{}\hfil$\crcr\noalign{\kern1.5pt}}}}}
\usepackage{bold-extra}
\usepackage{bm}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage[colorlinks = true,
            linkcolor = black,
            urlcolor  = blue,
            citecolor = black,
            anchorcolor = black]{hyperref}
\usepackage{epstopdf}
\usepackage{braket}
\usepackage{upgreek}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{subcaption}
\usepackage{amssymb,latexsym,amsmath,gensymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{float}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{url}
\usepackage{array}
\newcolumntype{C}{>{$\displaystyle} c <{$}}
\usepackage{tikz, calc}
\usetikzlibrary{shapes.geometric, arrows, calc}
\tikzstyle{norm} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{arrow} = [thick, ->, >=stealth]

\providecommand{\argmax}[1]{\underset{#1}{\mathrm{argmax}}\,}
\providecommand{\argmin}[1]{\underset{#1}{\mathrm{argmin}}\,}
\newcommand{\me}{\mathrm{e}}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}
\providecommand{\mb}[1]{\mathbf{#1}}
\providecommand{\mc}[1]{\mathcal{#1}}
\providecommand{\ro}{\mathbf{\mathfrak{r}}_o}
\providecommand{\so}{\mathbf{\hat{s}}_o}
\providecommand{\sd}{\mathbf{\hat{s}}_d}
\providecommand{\pp}{\mathbf{\hat{p}}}
\providecommand{\rb}{\mathbf{r}_b}
\providecommand{\rbm}{r_b^{\text{m}}}
\providecommand{\rd}{\mathbf{\mathfrak{r}}_d}
\providecommand{\mh}[1]{\mathbf{\hat{#1}}}
\providecommand{\mf}[1]{\mathfrak{#1}}
\providecommand{\ms}[1]{\mathsf{#1}}
\providecommand{\mbb}[1]{\mathbb{#1}}
\providecommand{\bs}[1]{\boldsymbol{#1}}
\providecommand{\tv}{\texttt{v}}
\providecommand{\tx}[1]{\text{#1}}
\providecommand{\tb}[1]{\textbf{#1}}
\providecommand{\ttt}[1]{\texttt{#1}}
\providecommand{\bv}{\bs{\nu}}
\providecommand{\bp}{\bs{\rho}}
\providecommand{\p}{\mh{p}}
\providecommand{\lmsum}{\sum_{l=0}^\infty\sum_{m=-l}^{l}}
\providecommand{\intr}[1]{\int_{\mbb{R}^{#1}}}
\providecommand{\ints}[1]{\int_{\mbb{S}^{#1}}}
\providecommand{\intinf}{\int_{-\infty}^{\infty}}
\let\originalleft\left
\let\originalright\right
\renewcommand{\left}{\mathopen{}\mathclose\bgroup\originalleft}
\renewcommand{\right}{\aftergroup\egroup\originalright}

\providecommand{\fig}[4]{
  % filename, width, caption, label
\begin{figure}[H]
 \captionsetup{width=1.0\linewidth}
 \centering
 \includegraphics[width = #2\textwidth]{#1}
 \caption{#3}
 \label{fig:#4}
\end{figure}
}

\makeatletter
\renewcommand*\env@matrix[1][*\c@MaxMatrixCols c]{%
  \hskip -\arraycolsep
  \let\@ifnextchar\new@ifnextchar
  \array{#1}}
\makeatother

\newcommand{\tensor}[1]{\overset{\text{\tiny$\leftrightarrow$}}{\mb{#1}}}
\newcommand{\tunderbrace}[2]{\underbrace{#1}_{\textstyle#2}}
\providecommand{\figs}[7]{
  % filename1, filename2, caption1, caption2, label1, label2, shift
\begin{figure}[H]
\centering
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#1}
  \captionsetup{justification=justified, singlelinecheck=true}
  \caption{#3}
  \label{fig:#5}
\end{minipage}
\hspace{2em}
\begin{minipage}[b]{.45\textwidth}
  \centering
  \includegraphics[width=1.0\linewidth]{#2}
  \vspace{#7em}
  \captionsetup{justification=justified}
  \caption{#4}
  \label{fig:#6}
\end{minipage}
\end{figure}
}
\makeatletter

\providecommand{\code}[1]{
\begin{center}
\lstinputlisting{#1}
\end{center}
}

\newcommand{\crefrangeconjunction}{--}
%%%%%%%%%%%
% Spacing %
%%%%%%%%%%%
% Margins
\usepackage[
top    = 1.5cm,
bottom = 1.5cm,
left   = 1.5cm,
right  = 1.5cm]{geometry}

% Indents, paragraph space
%\usepackage{parskip}
\setlength{\parskip}{1.5ex}

% Section spacing
\usepackage{titlesec}
\titlespacing*{\title}
{0pt}{0ex}{0ex}
\titlespacing*{\section}
{0pt}{0ex}{0ex}
\titlespacing*{\subsection}
{0pt}{0ex}{0ex}
\titlespacing*{\subsubsection}
{0pt}{0ex}{0ex}

% Line spacing
\linespread{1.1}

%%%%%%%%%%%%
% Document %
%%%%%%%%%%%%
\begin{document}
\title{\vspace{-2.5em} Optimizing angular diSPIM measurements\vspace{-1em}} % \author{Talon Chandler, Min Guo, Hari
  % Shroff, Rudolf Oldenbourg, Patrick La Rivi\`ere}
\date{\vspace{-3em}\today\vspace{-1em}}
\maketitle

\section{Introduction}
Our goal in these notes is to find optimal angular sampling schemes for the
diSPIM with light-sheet tilting. We review the continuous angular forward model,
find the sampling schemes available to us, choose an objective function to
optimize, then search through the sampling schemes.

\section{Continuous forward model}
We briefly review the continuous angular forward model---see the
\href{https://github.com/talonchandler/polharmonic/blob/master/notes/2018-12-10-angular-svd/report/report.pdf}{previous
  notes} for a more detailed discussion. We can write the
continuous-to-continuous forward model in the form
\begin{align}
  \mb{g}_c = \mc{H}_{cc}\mb{f}_c,\label{eq:cc}
\end{align}
where $\mb{f}_c \in \mbb{L}_2(\mbb{S}^2)$ is the angular dipole density,
$\mb{g}_c \in \mbb{L}_2(\mbb{S}^2\times \mbb{S}^1)$ is the continuous irradiance
data taken by varying the detection optical axis (the $\mbb{S}^2$ dimension) and
the illumination polarizer (the $\mbb{S}^1$ dimension), and $\mc{H}_{cc}$ is a
linear continuous-to-continuous Hilbert-space operator between these spaces.

If we choose a delta-function basis for both object and data space then we can
rewrite Eq. \eqref{eq:cc} as 
\begin{align}
  g(\sd, \pp) = \ints{2}d\so\, h(\sd,\pp;\so)f(\so),\label{eq:fwd}
\end{align}
where $\so \in \mbb{S}^2$ is the object angular coordinate, $\sd \in \mbb{S}^2$
is the detection optical axis coordinate, $\pp \in \mbb{S}^1$ is the polarizer
coordinate, and $h(\sd,\pp;\so)$ is the kernel of the integral transform. We can
separate the kernel into an excitation and detection part since these processes
are incoherent
\begin{align}
  h(\sd, \pp;\so) = h^{\text{exc}}(\pp;\so)h^{\text{det}}(\sd;\so). \label{eq:kern}
\end{align}
The excitation kernel is straightforward---the squared dot product of the
polarizer orientation $\pp$ with the object angular coordinate $\so$
\begin{align}
  h^{\text{exc}}(\pp\cdot\so) \propto (\pp\cdot\so)^2. \label{eq:exckern}
\end{align}
The detection kernel requires more care since we are collecting light over the
NA of the objective. In the previous notes we started with the power along a
single ray then integrated over the aperture to find that the detection kernel is
\begin{align}
  h^{\text{det}}(\sd\cdot\so) \propto 2(1 - \sqrt{\cos\alpha})P_0(\sd\cdot\so) + \frac{1}{5}\left[\frac{2}{5} + \frac{1}{10}\sqrt{\cos\alpha}(-7+3\cos\alpha)\right]P_2(\sd\cdot\so),
\end{align}
where $\text{NA} = n_o\sin\alpha$, and $P_{\ell}(x)$ are the Legendre
polynomials.

We can rewrite the mapping in Eq. \eqref{eq:cc} in an object-space spherical
harmonic basis as
\begin{align}
  g(\sd, \pp) = \sum_{\ell m}H_\ell^m(\sd, \pp)F_\ell^m,\label{eq:fwdsh}
\end{align}
where $F_\ell^m$ is the angular dipole spectrum given by 
\begin{align}
  F_\ell^m = \int_{\mbb{S}^2}d\so\, f(\so)Y_\ell^{m*}(\so),
\end{align}
and $H_\ell^m(\sd, \pp)$ is the kernel in this basis given by
\begin{align}
  H_\ell^m(\sd, \pp) &= \sum_{\ell' m'}\sum_{\ell'' m''}(-1)^m G_{\ell,\ell',\ell''}^{-m,m',m''}H_{\ell'}^{m',\text{exc}}(\pp)H_{\ell''}^{m'',\text{det}}(\sd),\label{eq:triple4}\\
    H_{\ell'}^{m',\text{exc}}(\pp) &\propto Y_{\ell'}^{m'*}(\pp)\left[\delta_{0\ell'} + \frac{2}{5}\delta_{2,\ell'}\right],\label{eq:dipexc}\\
    H_{\ell''}^{m'',\text{det}}(\sd) &\propto \frac{Y_{\ell''}^{m''*}(\sd)}{1 - \cos\alpha}\left[(1 - \sqrt{\cos\alpha})\delta_{0,\ell''} + \left(\frac{2}{5} + \frac{1}{10}\sqrt{\cos\alpha}(-7+3\cos\alpha)\right)\delta_{2,\ell''}\right].\label{eq:dipdet}
\end{align}
with the Gaunt coefficients given by 
\begin{align}
  G_{\ell,\ell',\ell''}^{m,m',m''} = \ints{2}d\so\, Y_\ell^{m}(\so)Y_{\ell'}^{m'}(\so)Y_{\ell''}^{m''}(\so).
\end{align}

It's useful to compare Eq. \eqref{eq:fwd} to Eq. \eqref{eq:fwdsh}---in a delta
function basis we need to compute an angular integral, while in the object-space
spherical harmonics basis the kernel has a finite number of terms so we can
compute it exactly.

It's also useful to compare Eq. \eqref{eq:kern} to Eq. \eqref{eq:triple4}---in a
delta function basis the kernel is the product of the excitation and detection
kernels, while in the object-space spherical harmonic basis the kernel is a
generalized convolution of the excitation and detection kernels.

\section{Sampling schemes}
We can write the sampling operation as a continuous-to-discrete Hilbert space
operator $\mc{D}_w$
\begin{align}
  \mb{g}_d = \mc{D}_w\mb{g}_c = \mc{D}_w\mc{H}_{cc}\mb{f}_c. 
\end{align}
We can write the sampling operation in a delta function basis as
\begin{align}
  g_m = \int_{\mbb{S}^2}d\sd\int_{\mbb{S}^1}d\pp\, w_m(\sd,\pp)g(\sd,\pp).
\end{align}
where $g_m$ is the $m$th discrete measurement and $w_m(\sd,\pp)$ is the $m$th
sampling aperture. For now we will restrict ourselves to delta-function sampling
apertures
\begin{align}
  w_m(\sd,\pp) = \delta(\mh{s}_{d,m} - \sd)\delta(\mh{p}_{m} - \pp),
\end{align}
where $\mh{s}_{d,m}$ and $\mh{p}_{m}$ are the $m$th sampling points. This
restriction means that we will leave the polarizer setting and tilt angle fixed
during each volume acquisition. We could consider scanning the polarizer
setting and tilt angle in the future for faster acquisitions, but I think these
sampling schemes will always decrease ``angular SNR''.

So far we've formulated everything quite generally and haven't made any
reference to the diSPIM geometry. Now we find the constraints that the diSPIM
geometry places on the values of $\mh{s}_{d,m}$ and $\mh{p}_{m}$. If we define
the $\mh{x}$ and $\mh{z}$ axes as the optical axes of the two objective then we
have the following sampling constraint
\begin{align}
\mh{s}_{d,m} \in \{\mh{x}, \mh{z}\}.  \label{eq:first}
\end{align}
In words, this constraint says that we can only detect along one of two discrete
and orthogonal axes.

Next, we constrain the polarizer sampling points $\mh{p}_m$. We introduce a
variable $\mh{s}_{e,m}$ to denote the excitation optical axis for the $m$th
measurement. This variable does not affect the forward model (we've formulated
in terms of $\pp$ only on the excitation side), but it will allow us to
conveniently constrain $\pp$. First, we constrain the excitation optical axis to
be perpendicular to the detection optical axis
\begin{align}
  \mh{s}_{e,m} \cdot \mh{s}_{d,m} = 0.
\end{align}
This is a fairly loose constraint, but it ensures that the light sheet is in
focus across the field of view.

Next, we constrain the excitation optical axis to within the available tilt
angles of the instrument. If we use $\mc{R}$ to denote an ``axis swapping
operator'' ($\mc{R}(\mh{z}) = \mh{x}$ and $\mc{R}(\mh{x}) = \mh{z}$), then the tilting
constraint can be written as
\begin{align}
  \mh{s}_{e,m} \cdot \mc{R}(\mh{s}_{d,m}) \leq \Delta,
\end{align}
where $\Delta$ is the maximum tilt angle. For the current instrument $\Delta = 15^{\circ} = \pi/12$.

Finally, we constrain the excitation polarization $\pp_m$ to be perpendicular to
the excitation optical axis
\begin{align}
  \mh{s}_{e,m} \cdot \pp_m = 0.  \label{eq:final}
\end{align}
For convenience we will denote individual sampling points with $\mf{s}_m$ =
($\pp_m$, $\mh{s}_{d,m}$), and we will denote the set of all sampling points
$\mf{s}_m$ that satisfy Eqs. \eqref{eq:first}--\eqref{eq:final} by $\mc{S}$.
With this notation we can denote a complete sampling scheme with $M$ samples
by $\mf{s} \in \mc{S}^M$ ($M$ copies of the set $\mc{S}^M$).

\section{Objective functions}
Our goal is to choose an $M$-sample sampling scheme $\mf{s} \in \mc{S}^M$ that
is optimal in some sense. We propose that we minimize the $\mbb{L}_2$ condition
number of the continuous-to-discrete imaging operator---the ratio of the largest
and smallest singular values of the matrix. We can interpret the condition
number as the ``invertibility'' of the matrix.

Optimizing the condition number will generate designs that are easily inverted,
but the designs may estimate a small number of uninteresting parameters. For
this reason we impose a rank constraint on the matrix. We require that
\begin{align}
  \text{rank}(\mc{D}_{\mf{s}}\mc{H}_{cc}\mc{P}_{\ell_c = 2}) \geq 6, \label{eq:constraint}
\end{align}
where $\mc{P}_{\ell_c=2}$ is a projection operator onto the $\ell=0$ and
$\ell=2$ subspace. For convenience we will denote the set of sampling schemes
that satisfy the constraint in Eq. \eqref{eq:constraint} by
$\mc{Z} = \{\mf{s}\, |\, \text{rank}(\mc{D}_{\mf{s}}\mc{H}_{cc}\mc{P}_{\ell_c =
  2}) \geq 6\}$. An optimal $M$-sample sampling scheme will be the solution of the optimization problem 
\begin{align}
  \mf{s}^* = \argmin{\mf{s} \in \mc{S}^M \cap \mc{Z}}\kappa(\mc{D}_{\mf{s}}\mc{H}_{cc}), \label{eq:notimekappa}
\end{align}
where $\kappa(A)$ denotes the $\mbb{L}_2$ condition number of $A$.

Earlier we mainly used the Schatten 2-norm of the continuous-discrete imaging
operator---the root sum of squares of the singular values---as the objective
function. This objective function resulted in useful designs, but the last
singular value was usually small which led us to optimizing the condition
number. We can interpret the Schatten 2-norm designs as the ``learn as much as
we can about all parameters'' designs and the condition number designs as the
``no parameter left behind'' designs.

% Schatten 2-norm of the
% continuous-discrete imaging operator
% \begin{align}
%   \mf{s}^* = \argmax{\mf{s} \in \mc{S}^M}||\mc{D}_{\mf{s}}\mc{H}_{cc}||_2. \label{eq:notime}
% \end{align}
% The Schatten $p$-norm is defined as the $\mbb{L}_p$ norm of the singular values
% of the operator, so the Schatten $2$-norm is a reasonable measure of the amount
% of information that the imaging system can pass. It favors sampling schemes that
% have many large singular values, so it will prefer sampling schemes that measure
% many orthogonal parameters of the object that are insensitive to noise.

% It's useful to interpret the Schatten $2$-norm by analyzing a spatial imaging
% system. The ``effective OTF'' that appears in many fluorescence microscopy
% papers (including Hari and Yicong's 2018 SIM review) is a specific example of a
% continuous singular value spectrum. We can calculate the Schatten $2$-norm by
% taking the effective OTF, squaring its value at each spatial frequency,
% integrating over the spatial frequencies, then taking the square root. The
% Schatten $2$-norm will be largest for imaging systems that have many singular
% values (a high spatial frequency bandwidth) that are large (we prefer effective
% OTFs that are large and constant over the bandwidth).

% Eq. \eqref{eq:notime} will optimize the number and size of the singular values
% without considering how interpretable the corresponding parameters are. One of
% our motivations for adding light-sheet tilting was to measure rotationally
% invariant parameters of the object which will be much more interpretable than
% rotationally variant parameters. To achieve this objective we need to measure at
% least the $\ell=0$ and $\ell = 2$ rotationally invariant subspaces---a
% 6-dimensional subspace of $\mbb{L}_2(\mbb{S}^2)$. This will require at least 6
% measurements. To ensure that a sampling scheme has measured this rotationally
% invariant subspace, we require that
% \begin{align}
%   \text{rank}(\mc{D}_{\mf{s}}\mc{H}_{cc}\mc{P}_{\ell_c = 2}) \geq 6, \label{eq:constraint}
% \end{align}
% where $\mc{P}_{\ell_c=2}$ is a projection operator onto the $\ell=0$ and
% $\ell=2$ subspace. For convenience we will denote the set of sampling schemes
% that satisfy the constraint in Eq. \eqref{eq:constraint} by
% $\mc{Z} = \{\mf{s}\, |\, \text{rank}(\mc{D}_{\mf{s}}\mc{H}_{cc}\mc{P}_{\ell_c = 2})
% \geq 6\}$.

% The solutions of Eq. \eqref{eq:notime} will not necessarily satisfy Eq.
% \eqref{eq:constraint}, and simply adding an extra constraint to Eq.
% \eqref{eq:constraint} will not ensure that we are optimally measuring the
% parameters in the subspace that we want to measure. Instead, we propose an
% alternative optimization problem
% \begin{align}
%     \mf{s}_{\ell_c =2}^* = \argmax{\mf{s} \in \mc{S}^M\cap\mc{Z}}||\mc{D}_{\mf{s}}\mc{H}_{cc}\mc{P}_{\ell_c = 2}||_2,
% \end{align}
% which ensures that the objective function is independent of singular values that
% measure parameters outside of the objective subspace. Note that the set
% $\mc{S}^M\cap\mc{Z}$ will definitely be empty for $M \leq 6$, and we are not sure
% what value of $M$ will yield solutions.

\section{Practical implementation details}
The optimization problem above lacks many of the details required for an
implementation. In this section I'll go through some of the practical work of
converting the abstract form into a finite-dimensional optimization problem.

First, we need to choose a basis and find matrix representations for the
operators $\mc{D}_{\mf{s}}$ and $\mc{H}_{cc}$. At first glance this seems
difficult---$\mc{D}_{\mf{s}}$ is a CD operator and $\mc{H}_{cc}$ is a CC
operator so they do not have finite-dimensional matrix representations
individually. Luckily, the combination $\mc{D}_{\mf{s}} \mc{H}_{cc}$ is a
band-limited compact operator, so it has a finite-dimensional representation as
a $15\times M$ matrix that we will call $\ms{H}$. The entries of this matrix are
given by
\begin{align}
  \ms{H}_{jm} = H_j(\mh{s}_{d,m}, \mh{p}_m),
\end{align}
where $j$ is a single index over the spherical harmonics, and Eq.
\ref{eq:triple4} shows us how to calculate the entries in terms
of the excitation and detection kernels.

Next, we need a matrix representation of $\mc{P}_{\ell_c=2}$. This is simple in
a spherical harmonics basis---we create a 15$\times$15 matrix called
$\ms{P}$, fill the first six diagonal elements with ones, and place zeroes
elsewhere
\begin{align}
  \ms{P}_{ij} = \delta_{ij}, \qquad i,j = 0,1,2,3,4,5.
\end{align}

% Next, we need a way to calculate the Schatten 2-norm efficiently. We could
% calculate the singular values of each $15\times 15$ matrix, but this would be
% slow. We can take a shortcut and replace the Schatten 2-norm with the Frobenius
% norm---the $\mbb{L}_2$ norm of the matrix entries. Note that this replacement is
% only valid for finite-dimensional representations of the operators.

Finally, we need an efficient set of scalar parameters to search through the set
of sampling schemes $\mc{S}^M$. We will parameterize each sample with a discrete
view index $v \in \{0, 1\}$ to indicate the detection axis, a tilt angle defined
from the nominal excitation axis $\delta \in [-\Delta, \Delta]$, and a polarizer
angle defined from the detection axis $\phi \in [0, 2\pi)$.

This parameter space is challenging to search because it is continuous in some
parameters and discrete in others. It's not clear if it's possible to take
derivatives with respect to the continuous parameters, so as a first pass we
will discretize the continuous parameters and use a search heuristic.
Discretization has the advantage of providing solutions that will be easier to
implement experimentally since we can constrain our solutions to sampling
schemes that are easy to align, but of course we can't guarantee that our
sampling scheme solutions will be globally optimal.

We will discretize the tilt angle into 3 angles
$\delta \in \{-\Delta, 0, \Delta\}$, and we will discretize the polarizer angle
into $P$ angles $\phi \in \{0,\pi/P,\dots,(P-1)\pi/P\}$. This means that a
brute-force approach will need to search the number of ways to choose $M$
elements from $2\,\,\text{views}\times 3\,\,\text{tilts}\times P\,\,\text{polarizer}$
options = $\binom{6P}{M}$.

My current implementation requires about 2 milliseconds per objective function
evaluation, so a brute force search is feasible for $P=6$ and $M=6$ since
$\binom{36}{6} \approx 2\e{6}$ evaluations will take about 2 hours.

\section{Results}\label{sec:results}
Figure 1--3 include several subfigures that we can use to compare the designs.

In the \textbf{top left} subfigure we show a schematic of the angular design.
Each arrow indicates a single polarized excitation direction and tilt angle. In
the current diSPIM design we do not use any excitation arm tilting, so all 6
arrows are centered on the optic axes (indicated by large black dots). The
colors of the arrows match the small colored squares in the \textbf{top center}
subfigure which indicate the point response functions for each measurement. The
radius of each point response function indicates the size of the signal measured
from a fluorophore oriented along that direction.

In the \textbf{top right} subfigure we show an set of single direction
reconstructions with the given design. We start by creating a uniformly oriented
phantom for each of the 250 directions in the positive octant of the sphere
indicated by \textcolor{green}{green} dots. Next, we simulate the imaging
process, simulate Poisson noise with SNR = 20, reconstruct the phantom, then
find maximum of the reconstructed ODF. We mark each reconstructed maximum with a
\textcolor{red}{red} dot and connect each phantom/reconstruction pair with a
black line.

In the \textbf{center} subfigure we show the singular value spectrum of the
imaging system with the singular vectors plotted on the $x$ axis. The singular
vectors are orthogonal functions, so they necessarily have positive
(\textcolor{red}{red}) and negative (\textcolor{blue}{blue}) components. In the
top-right of the figure we show the Schatten 2-norm of the design, the condition
number, and the rank.

In the \textbf{bottom} subfigure we show the singular value spectrum of the
imaging system after projecting onto the $\ell = 0, 2$ subspace with the associated
Schatten 2-norm, condition number, and rank.

\fig{../figures/custom-designs/3dispim.pdf}{1.0}{3-polarization diSPIM design.
  See the top of Section \ref{sec:results} for the general caption. The current design
  does not use any light sheet tilting, so the arrows in the schematic are
  centered on the black dots. Although the 3-polarization diSPIM design measures
  a 6-dimensional space (its rank is 6), the last singular value is very small
  (1/22 of the maximum singular value) so it is effectively lost in the noise.
  The 3-polarization diSPIM design only measures a 5-dimensional subspace of the
  $\ell = 0, 2$ subspace, so the reconstructions suffer from significant blind
  spots. The single direction reconstructions show that many directions are
  reconstructed with errors larger than 30 degrees.}{3dispim}

\fig{../figures/custom-designs/optimal-6pol.pdf}{1.0}{Optimal $M=6$, $P=3$
  tilting diSPIM design. See the top of Section \ref{sec:results} for the general
  caption. This design is the optimal $6$-measurement design when $3$ equally
  spaced polarizer settings and light-sheet tilting are available. First,
  collect the \textcolor{blue}{blue} samples by tilting the light sheet and
  cycling through the three polarizer settings. Next, collect the
  \textcolor{red}{red} samples from the other view by choosing the polarizer
  setting that points towards \textcolor{blue}{blue} excitation axis and tilting
  the excitation sheet both directions. Finally, collect the
  \textcolor{green}{green} sample by changing the polarizer setting and exciting
  along the optic axis. This design shows several advantages over our current
  design: the smallest singular value is 1/9 the size of the largest singular
  value, so we have a much better chance of recovering it from the noise; we
  measure components from all 6 dimensions of the $\ell=0, 2$ subspace; and the
  single direction reconstruction shows that we can reconstruct with errors less
  than 15 degrees.}{33dispim}

\fig{../figures/custom-designs/optimal-6pol-6choice.pdf}{1.0}{Optimal $P=6$
  tilting diSPIM design. See the top of Section \ref{sec:results} for the general
  caption. This design is the optimal $6$-measurement design when $6$ equally
  spaced polarizer settings and light-sheet tilting are available. Compared to
  the optimal $P=3$ design in Figure 2, the optimal $P=6$ design has a smallest
  singular value that is 1/7 the size of the largest singular value, so it is
  even less susceptible to corruption by noise. Interestingly, the singular
  direction reconstruction looks worse for the $P=6$ design than the $P=3$
  design. This isn't too troubling since we have optimized on the condition
  number, not the single direction reconstruction. Although this design has a
  slightly better condition number, it requires finer polarization alignment
  than the $P=3$ design, it requires tilting on both arms, and it does not
  provide a very large improvement. With these comments and the ``Min constraint''
  in mind, I'm happy to recommend the $P=3$ design.}{36dispim}

\end{document}